### 2025-03-26 paper/paper.qmd
more paper setup

---
title: "Your Paper Title"
subtitle: "So fancy!"
author:
    - Luca Carnegie
    - Chris Du
    - Russell Luchin
    - Randall Ni
abstract:
    This is the abstract of your paper. Write a brief summary of your research,
    including your objectives, methods, key results, and main conclusions.
    Keep it concise but informative.
keywords: [keyword1, keyword2, keyword3]
format:
    pdf:
        documentclass: article
        papersize: letter
        toc: true
        number-sections: true
        colorlinks: true
        fig-cap-location: bottom
        tbl-cap-location: top
        geometry: "margin=1in"  
bibliography: references.bib
execute:
    echo: false  # Hide code by default
    warning: false
    message: false
    jupyter: "python"
---

# Introduction and Motivation

In this section, explain what drives you to consider this topic and why it is important.

# Literature Review

Review similar research and establish how your work relates to existing studies.

# Research Question

Clearly state your primary research question(s) and hypotheses if applicable.

# Data

## Data Source, Collection, and Cleaning
## Data Exploration

```{python}

```

# Methodology

## Research Design

## Research Design

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nullam eu justo in nisl tincidunt tincidunt. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Phasellus lacinia, urna eget venenatis vestibulum, nisi metus feugiat nisi, vel faucibus magna sem quis tellus. Donec eget ullamcorper magna, non lacinia erat. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Maecenas id eleifend enim. Cras rhoncus, magna in tempus tempor, dolor magna facilisis nisi, at ultrices velit mi quis orci.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Vestibulum tortor quam, feugiat vitae, ultricies eget, tempor sit amet, ante. Donec eu libero sit amet quam egestas semper. Aenean ultricies mi vitae est. Mauris placerat eleifend leo. Quisque sit amet est et sapien ullamcorper pharetra. Vestibulum erat wisi, condimentum sed, commodo vitae, ornare sit amet, wisi. Aenean fermentum, elit eget tincidunt condimentum, eros ipsum rutrum orci, sagittis tempus lacus enim ac dui. Donec non enim in turpis pulvinar facilisis. Ut felis. Praesent dapibus, neque id cursus faucibus, tortor neque egestas augue, eu vulputate magna eros eu erat. Aliquam erat volutpat. Nam dui mi, tincidunt quis, accumsan porttitor, facilisis luctus, metus. Phasellus ultrices nulla quis nibh. Quisque a lectus. Donec consectetuer ligula vulputate sem tristique cursus. Nam nulla quam, gravida non, commodo a, sodales sit amet, nisi. Sed vitae justo eu ipsum maximus efficitur. Curabitur at dolor sed felis laoreet interdum vitae id mi. Fusce gravida pulvinar felis, a sagittis nisi sagittis non.

## Analytical Approach
## Model Specification

# Results and Interpretation

## Main Findings
## Analysis of Key Results

# Discussion

## Implications of Findings
## Limitations
## Robustness Checks
## Future Research Directions

# Conclusion

Summarize your key findings and their broader significance.

# References

### 2025-03-26 scripts/01-clean-data.py

@@ -1,0 +1,29 @@
+import pandas as pd
+
+# Load temperature data
+temperature_path = "temperatures-modern.xlsx"
+xls = pd.ExcelFile(temperature_path)
+df_temp = xls.parse('temperatures-modern')
+
+# Filter for January
+january_df = df_temp[df_temp['month'] == 'Jan']
+january_table_extended = january_df[['station_name', 'year', 'month', 'month_date', 'mean_temp_c']]
+
+# Display the cleaned January dataset
+print("January Temperatures Sample:")
+print(january_table_extended.head())
+
+# Load sakura data
+sakura_path = "sakura-modern.csv"
+sakura_df = pd.read_csv(sakura_path)
+
+# Display the sakura data
+print("\\nSakura Data Sample:")
+print(sakura_df.head())
+
+# Merge on station_name and year
+merged_df = pd.merge(sakura_df, january_table_extended, on=['station_name', 'year'], how='inner')
+
+# Display the merged dataset
+print("\\nMerged Data Sample:")
+print(merged_df.head())

### 2025-03-26 paper/paper.qmd
corrected author name from Chris Du to Chris Lu in paper files

@@
     - Luca Carnegie
-    - Chris Du
+    - Chris Lu
     - Russell Luchin

### 2025-03-26 paper/paper.qmd
add link to code and workflow in paper

@@
     jupyter: "python"
+thanks: "Code and workflow available at: "https://github.com/lcarnegie/sakura" 
 ---

### 2025-03-26 scripts/01-clean-data.py
add preamble comments to data simulation and cleaning scripts

@@ -1,3 +1,12 @@
+# ============= PREAMBLE =============
+# Last Edited By: Chris Lu
+# Date: March 25, 2025
+# Description: Script to clean and merge temperature and sakura blooming data
+#             - Loads temperature data and filters for January
+#             - Loads sakura (cherry blossom) data
+#             - Merges datasets by station name and year
+# ====================================
+
 import pandas as pd
 
 # Load temperature data

### 2025-03-26 scripts/03-explore-data.py

@@ -1,0 +1,5 @@
+# ============= PREAMBLE =============
+# Last Edited By: --
+# Date: --
+# Description: --
+# ====================================

### 2025-03-28 paper/paper.qmd
update paper formatting and references; adjust margins and improve layout

--- 2025-03-26 paper/paper.qmd
@@
     Keep it concise but informative.
-keywords: [keyword1, keyword2, keyword3]
 format:
@@
         tbl-cap-location: top
-        geometry: "margin=1in"  
+        geometry: "margin=1.5in"  
 bibliography: references.bib
@@
     jupyter: "python"
+nocite: "@*"
+thanks: "Code and workflow available at: https://github.com/lcarnegie/sakura" 
 ---

### 2025-03-28 scripts/01-clean-data.py
refactor data loading and processing in 01-clean-data.py; fixed historical data handling

@@ -8,30 +8,42 @@
 # ====================================
 
 import pandas as pd
+import numpy as np
 
-# Load temperature data
-temperature_path = "temperatures-modern.csv"
-df_temp = pd.read_csv(temperature_path)
+## LOAD DATA ##
 
-# Filter for January
-january_df = df_temp[df_temp['month'] == 'Jan']
-january_table_extended = january_df[['station_name', 'year', 'month', 'month_date', 'mean_temp_c']]
+df_modern_temp = pd.read_csv("data/raw data/cookson_cleaned_csvs/temperatures-modern.csv") # Load modern temperature
+df_modern_blooming = pd.read_csv("data/raw data/cookson_cleaned_csvs/sakura-modern.csv") # Load modern blooming data
+df_historical_temp = pd.read_csv("data/raw data/cookson_cleaned_csvs/sakura-historical.csv") # Load historical temperature data
 
-# Display the cleaned January dataset
-print("January Temperatures Sample:")
-print(january_table_extended.head())
+print("Modern Temp data\n" + str(df_modern_temp.head())) # Print first 5 rows of modern temperature data
+print()
+print("Modern bloom data\n" + str(df_modern_blooming.head())) # Print first 5 rows of modern blooming data
+print()
+print("Historical temp/bloom data\n" + str(df_historical_temp.head())) # Print first 5 rows of historical temperature data
 
-# Load sakura data
-sakura_path = "sakura-modern.csv"
-sakura_df = pd.read_csv(sakura_path)
+## HISTORICAL DATA ##
 
-# Display the sakura data
-print("\\nSakura Data Sample:")
-print(sakura_df.head())
+# drop unnecessary columns
+df_historical_temp = df_historical_temp.drop(columns=["study_source", "flower_source_name", "flower_source"])
 
-# Merge on station_name and year
-merged_df = pd.merge(sakura_df, january_table_extended, on=['station_name', 'year'], how='inner')
+# Merge reconstructed temp with observed temp into one col (prioritize observed)
+# If observed temp is NaN, fill with reconstructed temp
+df_historical_temp["temp_march"] = df_historical_temp["temp_c_obs"].combine_first(df_historical_temp["temp_c_recon"])
 
-# Display the merged dataset
-print("\\nMerged Data Sample:")
-print(merged_df.head())
+# convert day of year (where not NaN) to integer
[…]

### 2025-03-28 paper/paper.qmd
update paper margins and enhance bibliography entries; clarify reference details

--- 2025-03-26 paper/paper.qmd
@@
         tbl-cap-location: top
-        geometry: "margin=1in"  
+        geometry: "margin=1in" # print margins (change if wanting to)
 bibliography: references.bib
@@
     jupyter: "python"
+nocite: "@*" # display all references by default
+thanks: "Code and workflow available at: https://github.com/lcarnegie/sakura" 
 ---

### 2025-03-28 scripts/00-simulate-data.py
add preamble comments and enhance data simulation script; improve historical data transformation in cleaning script

@@ -1,5 +1,121 @@
 # ============= PREAMBLE =============
-# Last Edited By: --
-# Date: --
-# Description: --
+# Last Edited By: Luca Carnegie
+# Date: March 28, 2025
+# Description: 
+#       - Simulation script that generates synthetic cherry blossom 
+#         flowering data for Kyoto from 800 to 2024 CE, including dates 
+#         and emperatures. 
+#       - Creates visualizations showing historical flowering patterns 
+#         and climate relationships.
 # ====================================
+
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+from datetime import datetime, timedelta
+import random
+
+# Set random seed for reproducibility
+np.random.seed(42)
+
+# Create a function to generate simulated data
+def generate_kyoto_data(start_year=800, end_year=2024):
+    # Lists to store our data
+    years = []
+    dates = []
+    day_of_years = []
+    march_temps = []
+    
+    # Generate data for each year
+    for year in range(start_year, end_year + 1):
+        # Skip some years to simulate missing data
+        if random.random() < 0.1:  # 10% chance to skip a year
+            continue
+            
+        # Generate date of flowering
+        # Earlier dates in recent years to simulate climate change effect
+        base_day = 111  # Day of year from example (April 20)
+        
+        # Add some climate change effect - earlier blooming in recent years
+        climate_effect = int((year - start_year) / 100)  # Roughly 1 day earlier per century
+        
+        # Add some random variation
+        random_variation = np.random.normal(0, 7)  # Standard deviation of 7 days
+        
+        day_of_year = max(60, base_day - climate_effect - random_variation)  # Ensure not before March 1
+        
+        # Convert day of year to date
+        date = datetime(year, 1, 1) + timedelta(days=int(day_of_year) - 1)
+        date_str = f"{year}-{date.month:02d}-{date.day:02d}"
+        
+        # Generate March temperature
+        # Baseline temperature with some warming trend and random variation
+        base_temp = 12.0  # Base temperature for ancient records
+        warming_trend = 0.01 * (year - start_year)  # About 1°C per 100 years
[…]

### 2025-03-28 scripts/01-clean-data.py

@@ -22,7 +22,7 @@ print("Modern bloom data\n" + str(df_modern_blooming.head())) # Print first 5 ro
 print()
 print("Historical temp/bloom data\n" + str(df_historical_temp.head())) # Print first 5 rows of historical temperature data
 
-## HISTORICAL DATA ##
+## TRANSFORM HISTORICAL DATA ##
 
 # drop unnecessary columns
 df_historical_temp = df_historical_temp.drop(columns=["study_source", "flower_source_name", "flower_source"])
@@ -46,4 +46,15 @@ df_historical_temp = df_historical_temp[["year", "flower_date", "flower_doy", "t
 # cut off to 1952 and earlier (the rest we have modern data for)
 df_historical_temp = df_historical_temp[df_historical_temp["year"] <= 1952]
 
+## TRANSFORM MODERN DATA ##
 
+### Filter modern temp data
+
+### Filter modern flowering data 
+
+### Merge modern temp and flowering data
+
+## MERGE DATASETS ##
+
+## SAVE DATA ##
+# Save cleaned data to CSV files

### 2025-03-28 scripts/01-clean-data.py
finish cleaning data

@@ -12,9 +12,9 @@ import numpy as np
 
 ## LOAD DATA ##
 
-df_modern_temp = pd.read_csv("data/raw data/cookson_cleaned_csvs/temperatures-modern.csv") # Load modern temperature
-df_modern_blooming = pd.read_csv("data/raw data/cookson_cleaned_csvs/sakura-modern.csv") # Load modern blooming data
-df_historical_temp = pd.read_csv("data/raw data/cookson_cleaned_csvs/sakura-historical.csv") # Load historical temperature data
+df_modern_temp = pd.read_csv("data/raw_data/cookson_cleaned_csvs/temperatures-modern.csv") # Load modern temperature
+df_modern_blooming = pd.read_csv("data/raw_data/cookson_cleaned_csvs/sakura-modern.csv") # Load modern blooming data
+df_historical_temp = pd.read_csv("data/raw_data/cookson_cleaned_csvs/sakura-historical.csv") # Load historical temperature data
 
 print("Modern Temp data\n" + str(df_modern_temp.head())) # Print first 5 rows of modern temperature data
 print()
@@ -23,9 +23,6 @@ print()
 print("Historical temp/bloom data\n" + str(df_historical_temp.head())) # Print first 5 rows of historical temperature data
 
 ## TRANSFORM HISTORICAL DATA ##
-
-# drop unnecessary columns
-df_historical_temp = df_historical_temp.drop(columns=["study_source", "flower_source_name", "flower_source"])
 
 # Merge reconstructed temp with observed temp into one col (prioritize observed)
 # If observed temp is NaN, fill with reconstructed temp
@@ -40,21 +37,40 @@ df_historical_temp["temp_march"] = df_historical_temp["temp_march"].astype("Floa
 df_historical_temp["temp_c_obs"] = df_historical_temp["temp_c_obs"].astype("Float64")
 df_historical_temp["temp_c_recon"] = df_historical_temp["temp_c_recon"].astype("Float64")
 
+# drop unnecessary columns
+df_historical_temp = df_historical_temp.drop(columns=["study_source", "flower_source_name", "flower_source", "temp_c_obs", "temp_c_recon"])
+
 # reorder columns
-df_historical_temp = df_historical_temp[["year", "flower_date", "flower_doy", "temp_march", "temp_c_obs", "temp_c_recon"]]
+df_historical_temp = df_historical_temp[["year", "flower_date", "flower_doy", "temp_march"]]
+
[…]

### 2025-03-28 scripts/00-simulate-data.py
enhance data simulation and testing scripts; add preamble comments and implement comprehensive data validation tests

@@ -4,7 +4,7 @@
 # Description: 
 #       - Simulation script that generates synthetic cherry blossom 
 #         flowering data for Kyoto from 800 to 2024 CE, including dates 
-#         and emperatures. 
+#         and temperatures. 
 #       - Creates visualizations showing historical flowering patterns 
 #         and climate relationships.
 # ====================================

### 2025-03-28 scripts/02-test-data.py

@@ -1,5 +1,119 @@
 # ============= PREAMBLE =============
-# Last Edited By: --
-# Date: --
-# Description: --
+# Last Edited By: Luca Carnegie
+# Date: March 28, 2025
+# Description: 
+#        - Tests clean data script by loading the cleaned data and checking for
+#           missing values, duplicates, and basic statistics.
 # ====================================
+
+import os
+import pandas as pd
+import numpy as np
+import pytest
+from datetime import datetime
+
+# Path to the data file (assuming it's in a relative location to the script)
+DATA_PATH = os.path.join(os.path.dirname(os.path.dirname(__file__)), 
+                         'data', 'analysis_data', 'flowering_data.csv')
+
+def test_file_exists():
+    """Test that the data file exists."""
+    assert os.path.exists(DATA_PATH), f"Data file not found at {DATA_PATH}"
+
+def test_data_can_be_loaded():
+    """Test that the data can be loaded correctly."""
+    df = pd.read_csv(DATA_PATH)
+    assert isinstance(df, pd.DataFrame)
+    assert len(df) > 0, "DataFrame is empty"
+
+def test_schema():
+    """Test that the data has the expected columns."""
+    df = pd.read_csv(DATA_PATH)
+    expected_columns = ['year', 'flower_date', 'flower_doy', 'avg_temp_march_c']
+    assert all(col in df.columns for col in expected_columns), "Missing expected columns"
+
+def test_data_types():
+    """Test that the columns have the expected data types."""
+    df = pd.read_csv(DATA_PATH)
+    
+    # Test year is numeric
+    assert pd.api.types.is_numeric_dtype(df['year']), "Year should be numeric"
+    
+    # Test flower_date is string when not NaN
+    valid_dates = df['flower_date'].dropna()
+    if len(valid_dates) > 0:
+        assert pd.api.types.is_string_dtype(valid_dates), "flower_date should be string"
+    
+    # Test flower_doy and avg_temp_march_c are numeric
+    numeric_cols = ['flower_doy', 'avg_temp_march_c']
+    for col in numeric_cols:
+        assert pd.api.types.is_numeric_dtype(df[col]), f"{col} should be numeric"
+
+def test_value_ranges():
+    """Test that values are within expected ranges."""
+    df = pd.read_csv(DATA_PATH)
+    
+    # Year range
+    assert df['year'].min() >= 800, "Year values are too low"
+    assert df['year'].max() <= 1300, "Year values are too high"
+    
[…]

### 2025-03-28 paper/paper.qmd
Update paper.qmd
Added my portion of the lit review in.

@@
 
-Review similar research and establish how your work relates to existing studies.
+    The relationship between temperature and the flowering phenology of cherry blossoms (Prunus species, more commonly known as sakura) represents a unique merge of cultural heritage and biological science.
+For centuries, the timing of the sakura blooming has held profound cultural significance throughout East Asia, particularly in Japan, where the tradition of hanami (flower viewing) dates back to the Heian period.
+Beyond its cultural importance, the flowering phenology of sakura serves as a sensitive bioindicator of climate variability and change, with extensive historical records providing an exceptional opportunity to examine long-term temperature-phenology relationships (Aono & Kazui, 2008).
+
+    This review examines the extensive body of literature addressing the correlation between temperature and sakura flowering dates, integrating perspectives from plant biology, phenological modeling, climatology, and cultural studies.
+The biological mechanisms underlying temperature sensitivity in cherry blossoms involve complex interactions between winter chilling requirements and early spring temperatures (Allen et al., 2014).
+Concurrently, the cultural documentation of flowering dates through imperial court diaries, temple records, and festival calendars has created one of the longest continuous biological datasets in existence, with some records from Kyoto extending back to 812 CE (Aono & Kazui, 2008).
+
+    To begin with, Acevedo's thesis, "Sakura Storytellers" provides a rich cultural context for understanding cherry blossom phenology in Kyoto (2023).
+Her research highlights the 1200-year flowering record of Prunus jamasakura in Kyoto, describing it as "the longest and most complete phenological record in existence." These records, compiled from imperial court documents, poetry, and newspaper accounts of hanami celebrations, reveal that Kyoto's cherry trees have responded to major climate fluctuations throughout history, with progressive advancement in bloom dates occurring after approximately 1830.
+According to Aono and Kazui's analysis referenced in the thesis, cherry trees in Kyoto are currently experiencing warmer conditions than at any time in the past twelve centuries, resulting in average flowering dates that are seven days earlier in the 1971-2000 period compared to historical averages (2008).
+Beyond scientific significance, Acevedo explores how sakura in Kyoto have traditionally served as indicators of agricultural prosperity, with early-falling petals considered inauspicious for rice harvests, demonstrating their enduring role as both cultural and climatic storytellers in Japanese society.
+
+    Next, Xu et al.'s research provides significant physiological evidence supporting the temperature-flowering relationship in cherry trees (2023).
+Their study not only documented how temperature fluctuations caused 2-3 day shifts in flowering periods between 2016-2017, but also established varying cold tolerance rankings among cherry varieties that influence bloom timing.
+The authors revealed the underlying hormonal mechanisms—specifically changes in ABA, ZR, IAA, and GA3 levels—that regulate flower bud differentiation in response to temperature changes.
+This work strengthens our understanding of how temperature directly affects cherry blossom phenology through specific physiological pathways, offering valuable context for examining historical sakura flowering records in Kyoto and predicting how future climate patterns might alter traditional bloom timing.
+
+    In addition, Allen et al.'s research, "Modeling daily flowering probabilities" offers rigorous quantitative predictions about cherry blossom phenology under climate change in Japan (2014).
+They analyzed four cherry taxa (Prunus spachiana, P.
+× yedoensis, P.
+jamasakura, and P.
+lannesiana) using flexible Bayesian survival regression models applied to a 29-year dataset from Tama Forest Cherry Preservation Garden in Hachioji.
+Their approach incorporated both chilling requirements and heat accumulation mechanisms, finding that early-flowering species like P.
+spachiana were particularly sensitive to warming temperatures.
+The models predicted that all studied cherry taxa will flower approximately 30 days earlier by 2100 under the A1B emissions scenario, with 2-6 days greater uncertainty around mean flowering dates.
+The researchers find that sequential models (where trees require chilling before responding to warming) better predicted flowering times than simpler approaches.
+The study demonstrates that while cherry trees will continue to meet their chilling requirements under projected climate scenarios, accelerated heat accumulation will drive the dramatic advancement in flowering times
 

### 2025-03-28 paper/paper.qmd
changed how I wrote my reference so it actually use the reference.bib

@@
 For centuries, the timing of the sakura blooming has held profound cultural significance throughout East Asia, particularly in Japan, where the tradition of hanami (flower viewing) dates back to the Heian period.
-Beyond its cultural importance, the flowering phenology of sakura serves as a sensitive bioindicator of climate variability and change, with extensive historical records providing an exceptional opportunity to examine long-term temperature-phenology relationships (Aono & Kazui, 2008).
+Beyond its cultural importance, the flowering phenology of sakura serves as a sensitive bioindicator of climate variability and change, with extensive historical records providing an exceptional opportunity to examine long-term temperature-phenology relationships [@Aono2008].
 
@@
 The biological mechanisms underlying temperature sensitivity in cherry blossoms involve complex interactions between winter chilling requirements and early spring temperatures (Allen et al., 2014).
-Concurrently, the cultural documentation of flowering dates through imperial court diaries, temple records, and festival calendars has created one of the longest continuous biological datasets in existence, with some records from Kyoto extending back to 812 CE (Aono & Kazui, 2008).
+Concurrently, the cultural documentation of flowering dates through imperial court diaries, temple records, and festival calendars has created one of the longest continuous biological datasets in existence, with some records from Kyoto extending back to 812 CE [@Aono2008].
 
-    To begin with, Acevedo's thesis, "Sakura Storytellers" provides a rich cultural context for understanding cherry blossom phenology in Kyoto (2023).
+    To begin with, Acevedo's thesis, "Sakura Storytellers" provides a rich cultural context for understanding cherry blossom phenology in Kyoto [@Acevedo2023].
 Her research highlights the 1200-year flowering record of Prunus jamasakura in Kyoto, describing it as "the longest and most complete phenological record in existence." These records, compiled from imperial court documents, poetry, and newspaper accounts of hanami celebrations, reveal that Kyoto's cherry trees have responded to major climate fluctuations throughout history, with progressive advancement in bloom dates occurring after approximately 1830.
-According to Aono and Kazui's analysis referenced in the thesis, cherry trees in Kyoto are currently experiencing warmer conditions than at any time in the past twelve centuries, resulting in average flowering dates that are seven days earlier in the 1971-2000 period compared to historical averages (2008).
+According to Aono and Kazui's analysis referenced in the thesis, cherry trees in Kyoto are currently experiencing warmer conditions than at any time in the past twelve centuries, resulting in average flowering dates that are seven days earlier in the 1971-2000 period compared to historical averages [@Aono2008].
 Beyond scientific significance, Acevedo explores how sakura in Kyoto have traditionally served as indicators of agricultural prosperity, with early-falling petals considered inauspicious for rice harvests, demonstrating their enduring role as both cultural and climatic storytellers in Japanese society.
 
-    Next, Xu et al.'s research provides significant physiological evidence supporting the temperature-flowering relationship in cherry trees (2023).
+    Next, Xu et al.'s research provides significant physiological evidence supporting the temperature-flowering relationship in cherry trees [@Xu2023].
 Their study not only documented how temperature fluctuations caused 2-3 day shifts in flowering periods between 2016-2017, but also established varying cold tolerance rankings among cherry varieties that influence bloom timing.
@@
 
-    In addition, Allen et al.'s research, "Modeling daily flowering probabilities" offers rigorous quantitative predictions about cherry blossom phenology under climate change in Japan (2014).
+    In addition, Allen et al.'s research, "Modeling daily flowering probabilities" offers rigorous quantitative predictions about cherry blossom phenology under climate change in Japan [@Allen2014].
 They analyzed four cherry taxa (Prunus spachiana, P.
@@
 The researchers find that sequential models (where trees require chilling before responding to warming) better predicted flowering times than simpler approaches.
-The study demonstrates that while cherry trees will continue to meet their chilling requirements under projected climate scenarios, accelerated heat accumulation will drive the dramatic advancement in flowering times
+The study demonstrates that while cherry trees will continue to meet their chilling requirements under projected climate scenarios, accelerated heat accumulation will drive the dramatic advancement in flowering times.
 

### 2025-03-29 paper/paper.qmd
added first draft of intro/motivation and edited some citation

@@
 
-In this section, explain what drives you to consider this topic and why it is important.
+    The annual blooming of cherry blossoms (Prunus species, more commonly known as sakura) represents not only a culturally significant event across many East Asian countries but also serves as a valuable biological indicator of climate patterns and changes; particularly, the tradition of hanami (flower viewing) dates all the way back to the Heian period in Japan.
+This paper will examine the statistical relationship between temperature fluctuations and the timing of sakura flowering dates, a connection that has gained increased scientific attention as global climate change accelerates.
+Historical records of cherry blossom flowering, some dating back several centuries in Japan, provide an exceptionally rare long-term dataset for analyzing phenological responses to temperature variations [@Aono2008].
+By quantifying this relationship through robust statistical methods, we aim to contribute to the growing body of research on biological responses to climate change while also contextualizing the potential cultural impacts of shifting bloom times.
+The preservation of these iconic natural events holds both scientific value for understanding ecosystem responses to warming trends and cultural significance for communities where cherry blossom festivals represent important economic and social traditions.
 
@@
 
-    The relationship between temperature and the flowering phenology of cherry blossoms (Prunus species, more commonly known as sakura) represents a unique merge of cultural heritage and biological science.
-For centuries, the timing of the sakura blooming has held profound cultural significance throughout East Asia, particularly in Japan, where the tradition of hanami (flower viewing) dates back to the Heian period.
-Beyond its cultural importance, the flowering phenology of sakura serves as a sensitive bioindicator of climate variability and change, with extensive historical records providing an exceptional opportunity to examine long-term temperature-phenology relationships [@Aono2008].
-
-    This review examines the extensive body of literature addressing the correlation between temperature and sakura flowering dates, integrating perspectives from plant biology, phenological modeling, climatology, and cultural studies.
-The biological mechanisms underlying temperature sensitivity in cherry blossoms involve complex interactions between winter chilling requirements and early spring temperatures (Allen et al., 2014).
+    This literature review examines the extensive body of scholarly journals and articles addressing the correlation between temperature and sakura flowering dates, integrating perspectives from plant biology, phenological modeling, climatology, and cultural studies.
+The biological mechanisms underlying temperature sensitivity in cherry blossoms involve complex interactions between winter chilling requirements and early spring temperatures [@Allen2014].
 Concurrently, the cultural documentation of flowering dates through imperial court diaries, temple records, and festival calendars has created one of the longest continuous biological datasets in existence, with some records from Kyoto extending back to 812 CE [@Aono2008].

### 2025-03-29 scripts/01-clean-data.py
update data cleaning and testing scripts; enhance preamble comments, adjust year range validation, and add visualizations for average temperature and flowering date

@@ -1,7 +1,8 @@
 # ============= PREAMBLE =============
-# Last Edited By: Chris Lu
-# Date: March 25, 2025
-# Description: Script to clean and merge temperature and sakura blooming data
+# Last Edited By: Luca Carnegie
+# Date: March 28, 2025
+# Description: 
+#             - Script to clean and merge temperature and sakura blooming data
 #             - Loads temperature data and filters for January
 #             - Loads sakura (cherry blossom) data
 #             - Merges datasets by station name and year
@@ -43,7 +44,7 @@ df_historical_temp = df_historical_temp.drop(columns=["study_source", "flower_so
 # reorder columns
 df_historical_temp = df_historical_temp[["year", "flower_date", "flower_doy", "temp_march"]]
 
-#rename cols
+# rename cols
 df_historical_temp = df_historical_temp.rename(columns={"temp_march": "avg_temp_march_c"})
 
 # cut off to 1952 and earlier (the rest we have modern data for)
@@ -71,6 +72,10 @@ df_modern_bloom = df_modern_bloom.rename(columns={"mean_temp_c": "avg_temp_march
 
 df_flower = pd.concat([df_historical_bloom, df_modern_bloom], ignore_index=True)
 
+# change day of year to integer
+df_flower["flower_doy"] = df_flower["flower_doy"].astype("Int64")
+df_flower["flower_date"] = df_flower["flower_date"].fillna(pd.NA) # fill flower_date with <NA>
+
 ## SAVE DATA ##
 # Save cleaned data to CSV files
-df_flower.to_csv("data/analysis_data/flowering_data.csv", index=False) # Save cleaned data to CSV file
+df_flower.to_csv("data/analysis_data/clean_data.csv", index=False) # Save cleaned data to CSV file

### 2025-03-29 scripts/02-test-data.py

@@ -55,7 +55,7 @@ def test_value_ranges():
     
     # Year range
     assert df['year'].min() >= 800, "Year values are too low"
-    assert df['year'].max() <= 1300, "Year values are too high"
+    assert df['year'].max() <= 2024, "Year values are too high"
     
     # DOY range (day of year)
     valid_doy = df['flower_doy'].dropna()
@@ -117,3 +117,6 @@ def test_report_missing_values():
     
     # This isn't an assertion, just informational
     assert True
+
+if __name__ == "__main__":
+    pytest.main([__file__])

### 2025-03-29 scripts/03-explore-data.py

@@ -1,5 +1,57 @@
 # ============= PREAMBLE =============
-# Last Edited By: --
-# Date: --
-# Description: --
+# Last Edited By: Luca Carnegie
+# Date: March 28, 2025
+# Description: Script that creates some visualizations of the data
 # ====================================
+
+# Import libraries
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+import numpy as np
+
+# Read in data
+df = pd.read_csv('data/analysis_data/clean_data.csv')
+
+## AVERAGE MARCH TEMPERATURE ##
+
+# Histogram of average march temperature
+
+x = df['avg_temp_march_c']
+plt.figure(figsize=(10, 6))
+plt.hist(x, bins=20, color='blue', alpha=0.7)
+plt.title('Histogram of Average March Temperature in Kyoto, Japan')
+plt.xlabel('Average March Temperature (C)')
+plt.ylabel('Frequency')
+plt.grid(axis='y', alpha=0.75)
+plt.show()
+
+## FLOWERING DATE ##
+
+# Histogram of flowering date
+x = df['flower_doy']
+plt.figure(figsize=(10, 6))
+plt.hist(x, bins=np.arange(60, 125, 3), color='green', alpha=0.7)  # Changed bin size from 5 to 3
+plt.title('Histogram of Flowering Date in Kyoto, Japan')
+plt.xlabel('Month')
+plt.ylabel('Frequency')
+plt.grid(axis='y', alpha=0.75)
+
+# Convert DOY to month labels
+month_ticks = [60, 75, 90, 105, 120]  # DOY values
+month_labels = ['Mar 1', 'Mar 15', 'Apr 1', 'Apr 15', 'May 1']  # Corresponding month labels
+plt.xticks(month_ticks, month_labels, rotation=45)
+plt.show()
+
+## Scatterplot of average temperature vs flowering date
+y = df['flower_doy']
+x = df['avg_temp_march_c']
+plt.figure(figsize=(10, 6))
+plt.scatter(x, y, color='purple', alpha=0.5)
+plt.title('Scatterplot of Average March Temperature vs Flowering Date in Kyoto, Japan')
+plt.xlabel('Average March Temperature (C)')
+plt.ylabel('Flowering Date (DOY)')
+plt.grid(alpha=0.75)
+plt.show()
+
+

### 2025-03-29 paper/paper.qmd
minor citation change

@@
 
-    To begin with, Acevedo's thesis, "Sakura Storytellers" provides a rich cultural context for understanding cherry blossom phenology in Kyoto [@Acevedo2023].
+    To begin with, Acevedo's thesis, "Sakura Storytellers" provides a rich cultural context for understanding cherry blossom phenology in Kyoto [@Acevedo2021].
 Her research highlights the 1200-year flowering record of Prunus jamasakura in Kyoto, describing it as "the longest and most complete phenological record in existence." These records, compiled from imperial court documents, poetry, and newspaper accounts of hanami celebrations, reveal that Kyoto's cherry trees have responded to major climate fluctuations throughout history, with progressive advancement in bloom dates occurring after approximately 1830.

### 2025-03-29 scripts/01-clean-data.py
refactor data processing scripts; add censored column, save historic data, and remove exploration script

@@ -47,7 +47,13 @@ df_historical_temp = df_historical_temp[["year", "flower_date", "flower_doy", "t
 # rename cols
 df_historical_temp = df_historical_temp.rename(columns={"temp_march": "avg_temp_march_c"})
 
-# cut off to 1952 and earlier (the rest we have modern data for)
+# add column name for censored or not (1 for censored, 0 for not censored)
+df_historical_temp["censored"] = np.where(df_historical_temp["flower_date"].isna(), 1, 0)
+
+# save historical data to CSV file for one model
+df_historical_temp.to_csv("data/analysis_data/historic_bloom.csv", index=False) # Save historical data to CSV file
+
+# then cut off to 1952 and earlier (merge with modern data)
 df_historical_bloom = df_historical_temp[df_historical_temp["year"] <= 1952]
 
 ## TRANSFORM MODERN DATA ##
@@ -68,6 +74,9 @@ df_modern_bloom = df_modern_blooming.merge(df_modern_temp, how="inner", on=["yea
 ### rename cols
 df_modern_bloom = df_modern_bloom.rename(columns={"mean_temp_c": "avg_temp_march_c"})
 
+# add column name for censored or not (1 for censored, 0 for not censored)
+df_modern_bloom["censored"] = np.where(df_historical_temp["flower_date"].isna(), 1, 0)
+
 ## MERGE DATASETS ##
 
 df_flower = pd.concat([df_historical_bloom, df_modern_bloom], ignore_index=True)
@@ -78,4 +87,4 @@ df_flower["flower_date"] = df_flower["flower_date"].fillna(pd.NA) # fill flower_
 
 ## SAVE DATA ##
 # Save cleaned data to CSV files
-df_flower.to_csv("data/analysis_data/clean_data.csv", index=False) # Save cleaned data to CSV file
+df_flower.to_csv("data/analysis_data/merged_historic_modern_bloom.csv", index=False) # Save cleaned data to CSV file

### 2025-03-29 scripts/03-explore-data.py

@@ -1,57 +1,0 @@
-# ============= PREAMBLE =============
-# Last Edited By: Luca Carnegie
-# Date: March 28, 2025
-# Description: Script that creates some visualizations of the data
-# ====================================
-
-# Import libraries
-import pandas as pd
-import matplotlib.pyplot as plt
-import seaborn as sns
-import numpy as np
-
-# Read in data
-df = pd.read_csv('data/analysis_data/clean_data.csv')
-
-## AVERAGE MARCH TEMPERATURE ##
-
-# Histogram of average march temperature
-
-x = df['avg_temp_march_c']
-plt.figure(figsize=(10, 6))
-plt.hist(x, bins=20, color='blue', alpha=0.7)
-plt.title('Histogram of Average March Temperature in Kyoto, Japan')
-plt.xlabel('Average March Temperature (C)')
-plt.ylabel('Frequency')
-plt.grid(axis='y', alpha=0.75)
-plt.show()
-
-## FLOWERING DATE ##
-
-# Histogram of flowering date
-x = df['flower_doy']
-plt.figure(figsize=(10, 6))
-plt.hist(x, bins=np.arange(60, 125, 3), color='green', alpha=0.7)  # Changed bin size from 5 to 3
-plt.title('Histogram of Flowering Date in Kyoto, Japan')
-plt.xlabel('Month')
-plt.ylabel('Frequency')
-plt.grid(axis='y', alpha=0.75)
-
-# Convert DOY to month labels
-month_ticks = [60, 75, 90, 105, 120]  # DOY values
-month_labels = ['Mar 1', 'Mar 15', 'Apr 1', 'Apr 15', 'May 1']  # Corresponding month labels
-plt.xticks(month_ticks, month_labels, rotation=45)
-plt.show()
-
-## Scatterplot of average temperature vs flowering date
-y = df['flower_doy']
-x = df['avg_temp_march_c']
-plt.figure(figsize=(10, 6))
-plt.scatter(x, y, color='purple', alpha=0.5)
-plt.title('Scatterplot of Average March Temperature vs Flowering Date in Kyoto, Japan')
-plt.xlabel('Average March Temperature (C)')
-plt.ylabel('Flowering Date (DOY)')
-plt.grid(alpha=0.75)
-plt.show()
-
-

### 2025-03-29 scripts/04-model-data.py

@@ -1,5 +1,7 @@
-# ============= PREAMBLE =============
-# Last Edited By: --
-# Date: --
-# Description: --
-# ====================================
+
+
+
+
+# Start with the historic data only
+SakuraHistoric = load_data("sakura_historic_only") # Load the data
+SakuraHistoric.columns 

### 2025-03-31 paper/paper.qmd
remove unused model data script and refine introduction in paper

@@
     The annual blooming of cherry blossoms (Prunus species, more commonly known as sakura) represents not only a culturally significant event across many East Asian countries but also serves as a valuable biological indicator of climate patterns and changes; particularly, the tradition of hanami (flower viewing) dates all the way back to the Heian period in Japan.
-This paper will examine the statistical relationship between temperature fluctuations and the timing of sakura flowering dates, a connection that has gained increased scientific attention as global climate change accelerates.
+This paper will examine the relationship between temperature fluctuations and the timing of sakura flowering dates, a connection that has gained increased scientific attention as global climate change accelerates.
 Historical records of cherry blossom flowering, some dating back several centuries in Japan, provide an exceptionally rare long-term dataset for analyzing phenological responses to temperature variations [@Aono2008].

### 2025-03-31 scripts/04-model-data.py

@@ -1,7 +1,0 @@
-
-
-
-
-# Start with the historic data only
-SakuraHistoric = load_data("sakura_historic_only") # Load the data
-SakuraHistoric.columns 

### 2025-03-31 scripts/01-clean-data.py
re-cleaned wonky data

@@ -47,26 +47,23 @@ df_historical_temp = df_historical_temp[["year", "flower_date", "flower_doy", "t
 # rename cols
 df_historical_temp = df_historical_temp.rename(columns={"temp_march": "avg_temp_march_c"})
 
-# add column name for censored or not (1 for censored, 0 for not censored)
-df_historical_temp["censored"] = np.where(df_historical_temp["flower_date"].isna(), 1, 0)
-
 # save historical data to CSV file for one model
 df_historical_temp.to_csv("data/analysis_data/historic_bloom.csv", index=False) # Save historical data to CSV file
 
-# then cut off to 1952 and earlier (merge with modern data)
-df_historical_bloom = df_historical_temp[df_historical_temp["year"] <= 1952]
 
 ## TRANSFORM MODERN DATA ##
 
 ### Filter modern temp data to just Kyoto and March Temps (same as historical) 
 df_modern_temp = df_modern_temp[df_modern_temp["station_name"] == "Kyoto"]
 df_modern_temp = df_modern_temp[df_modern_temp["month"] == "Mar"]
-df_modern_temp = df_modern_temp[df_modern_temp["year"] >= 1953] # cut off to 1953 and later (the rest we have historical data for)
 df_modern_temp = df_modern_temp.drop(columns=["station_id", "station_name", "month", "month_date"])
 
 ### Filter modern flowering data to just Kyoto and March Temps 
 df_modern_blooming = df_modern_blooming[df_modern_blooming["station_name"] == "Kyoto"]
 df_modern_blooming = df_modern_blooming.drop(columns=["station_id", "station_name", "latitude", "longitude", "full_bloom_date", "full_bloom_doy"])
+
+# convert flower_doy to integer
+df_modern_blooming["flower_doy"] = df_modern_blooming["flower_doy"].astype("Int64")
 
 ### Merge modern temp and flowering data
 df_modern_bloom = df_modern_blooming.merge(df_modern_temp, how="inner", on=["year"])
@@ -74,17 +71,6 @@ df_modern_bloom = df_modern_blooming.merge(df_modern_temp, how="inner", on=["yea
 ### rename cols
 df_modern_bloom = df_modern_bloom.rename(columns={"mean_temp_c": "avg_temp_march_c"})
 
-# add column name for censored or not (1 for censored, 0 for not censored)
-df_modern_bloom["censored"] = np.where(df_historical_temp["flower_date"].isna(), 1, 0)
[…]

### 2025-03-31 paper/paper.qmd
Add new references for Japan Meteorological Agency and flowering data; rewrote rondales intro

@@
 
-# Introduction and Motivation
+# Introduction
 
-    The annual blooming of cherry blossoms (Prunus species, more commonly known as sakura) represents not only a culturally significant event across many East Asian countries but also serves as a valuable biological indicator of climate patterns and changes; particularly, the tradition of hanami (flower viewing) dates all the way back to the Heian period in Japan.
-This paper will examine the relationship between temperature fluctuations and the timing of sakura flowering dates, a connection that has gained increased scientific attention as global climate change accelerates.
-Historical records of cherry blossom flowering, some dating back several centuries in Japan, provide an exceptionally rare long-term dataset for analyzing phenological responses to temperature variations [@Aono2008].
-By quantifying this relationship through robust statistical methods, we aim to contribute to the growing body of research on biological responses to climate change while also contextualizing the potential cultural impacts of shifting bloom times.
-The preservation of these iconic natural events holds both scientific value for understanding ecosystem responses to warming trends and cultural significance for communities where cherry blossom festivals represent important economic and social traditions.
+Before the advent of modern technology, people understood themselves, their environments, and their cultures through the cycle of life and death.
+The Japanese tradition of *hanami* (loosely translated as 'flower-viewing') is an important cultural tradition that honors this perspective, being associated with the flowering birth of cherry blossom trees in the late spring.
+In the city of Kyoto in particular, local newspapers have recorded the dates of tree flowering since the late 19th century, but researchers like @Aono2010 have been able to estimate flowering dates as far back as the year 812 CE, by combing through historical records such as diaries, poetry and other written documents describing *hanami* parties throughout the ages.
+However, rising global temperatures create an anomaly, highlighting the need to understand the historical and modern links between temperature and cherry blossom bloom timing.
+Without it, cherished traditions like *hanami* may struggle to adapt to climate shifts, risking their disappearance as a unique Japanese custom.
+
+In this paper, we analyze the nearly 1200-year record compiled by @Aono2010 as well as modern data compiled by the Japanese Meteorological Agency (JMA) [@JMA_sakura] [@JMA_temp], compiled together by
+
+```
+
+This paper first focuses on quantifying a relationship between mean March temperature and the day of blooming of cherry blossom, using a mixture of historical data (@A)
+```
 
@@
 
-    This literature review examines the extensive body of scholarly journals and articles addressing the correlation between temperature and sakura flowering dates, integrating perspectives from plant biology, phenological modeling, climatology, and cultural studies.
-The biological mechanisms underlying temperature sensitivity in cherry blossoms involve complex interactions between winter chilling requirements and early spring temperatures [@Allen2014].
-Concurrently, the cultural documentation of flowering dates through imperial court diaries, temple records, and festival calendars has created one of the longest continuous biological datasets in existence, with some records from Kyoto extending back to 812 CE [@Aono2008].
+```         
+The relationship between sakura blooming date and other factors was first examined in depth.
+A notable amount of scholarship addresses the relationship between different variables and and the flowering dates of sakura, from humanistic cultural studies perspectives ranging to scientific plant biology, phenological modeling, and climatology studies.
 
-    To begin with, Acevedo's thesis, "Sakura Storytellers" provides a rich cultural context for understanding cherry blossom phenology in Kyoto [@Acevedo2021].
-Her research highlights the 1200-year flowering record of Prunus jamasakura in Kyoto, describing it as "the longest and most complete phenological record in existence." These records, compiled from imperial court documents, poetry, and newspaper accounts of hanami celebrations, reveal that Kyoto's cherry trees have responded to major climate fluctuations throughout history, with progressive advancement in bloom dates occurring after approximately 1830.
-According to Aono and Kazui's analysis referenced in the thesis, cherry trees in Kyoto are currently experiencing warmer conditions than at any time in the past twelve centuries, resulting in average flowering dates that are seven days earlier in the 1971-2000 period compared to historical averages [@Aono2008].
+@Acevedo2021 provides a rich cultural context for understanding cherry blossom phenology in Kyoto.
+Her research highlights a 1200-year flowering record of *Prunus jamasakura* in Kyoto, describing it as "the longest and most complete phenological record in existence." These records, compiled from imperial court documents, poetry, and newspaper accounts of *hanami* celebrations, reveal that Kyoto's cherry trees have responded to major climate fluctuations throughout history, with progressive advancement in bloom dates occurring after approximately 1830.
+This shift aligns with the end of the Little Ice Age and the onset of modern warming trends.
 Beyond scientific significance, Acevedo explores how sakura in Kyoto have traditionally served as indicators of agricultural prosperity, with early-falling petals considered inauspicious for rice harvests, demonstrating their enduring role as both cultural and climatic storytellers in Japanese society.
+
+According to @Aono2008, cherry trees in Kyoto are currently experiencing warmer conditions than at any time in the past twelve centuries, resulting in average flowering dates that are seven days earlier in the 1971-2000 period compared to historical averages.
 
@@
 
-    In addition, Allen et al.'s research, "Modeling daily flowering probabilities" offers rigorous quantitative predictions about cherry blossom phenology under climate change in Japan [@Allen2014].
+@Allen2014, "Modeling daily flowering probabilities" offers an approach to rigorous quantitative predictions about cherry blossom phenology under climate change in Japan.
 They analyzed four cherry taxa (Prunus spachiana, P.
@@
 lannesiana) using flexible Bayesian survival regression models applied to a 29-year dataset from Tama Forest Cherry Preservation Garden in Hachioji.
-Their approach incorporated both chilling requirements and heat accumulation mechanisms, finding that early-flowering species like P.
-spachiana were particularly sensitive to warming temperatures.
+Their approach incorporated both chilling requirements and heat accumulation mechanisms, finding that early-flowering species like *P.
+spachiana* were particularly sensitive to warming temperatures.
 The models predicted that all studied cherry taxa will flower approximately 30 days earlier by 2100 under the A1B emissions scenario, with 2-6 days greater uncertainty around mean flowering dates.
-The researchers find that sequential models (where trees require chilling before responding to warming) better predicted flowering times than simpler approaches.
+Allen et al.
+found that sequential models (where trees require chilling before responding to warming) better predicted flowering times than simpler approaches.
 The study demonstrates that while cherry trees will continue to meet their chilling requirements under projected climate scenarios, accelerated heat accumulation will drive the dramatic advancement in flowering times.
+```
 
@@
 Clearly state your primary research question(s) and hypotheses if applicable.
-
-# Data
-
-## Data Source, Collection, and Cleaning
-## Data Exploration
-
-```{python}
-
-```
 
@@
 ## Analytical Approach
+
+# Data
+
+## Data Source, Collection, and Cleaning
+
+## Data Exploration
+
+```{python}
+
+```
+
+# Modeling
+
 ## Model Specification
 
-# Results and Interpretation
-
-## Main Findings
-## Analysis of Key Results
+## Results
 
@@
 ## Implications of Findings
+
 ## Limitations
+
 ## Robustness Checks
+
 ## Future Research Directions

### 2025-03-31 paper/paper.qmd
Add initial draft of research paper on cherry blossom phenology and climate change

@@
 title: "Your Paper Title"
-subtitle: "So fancy!"
+subtitle: "Understanding Temperature's Role in Earlier Cherry Blossom Blooming Times in Kyoto, Japan"
 author:
@@
         tbl-cap-location: top
-        geometry: "margin=1in" # print margins (change if wanting to)
+        fontsize: "12"
+        geometry: "margin=1.2in" # print margins (change if wanting to)
 bibliography: references.bib
@@
 Without it, cherished traditions like *hanami* may struggle to adapt to climate shifts, risking their disappearance as a unique Japanese custom.
+To this end, work like @Allen2014's analyse and relatively modern blooming data over a 29-year period, but the key weakness of this is a lack of perspective.
+Given that their data only covers a small part of the entire recorded history of cherry blossom blooming, their analysis and subsequent model have a key weakness - they do not take into account large scale trends of blooming and their relationship to the change in climate that has happened in the last 150 years, which @Aono2010's data can illustrate.
 
-In this paper, we analyze the nearly 1200-year record compiled by @Aono2010 as well as modern data compiled by the Japanese Meteorological Agency (JMA) [@JMA_sakura] [@JMA_temp], compiled together by
+In this paper, we analyse and model the nearly 1200-year record compiled by @Aono2010 along modern data compiled by the Japanese Meteorological Agency (JMA) [@JMA_sakura] [@JMA_temp], collected together by @Cookson.
+In particular, we seek to answer the question of whether warmer temperatures lead to earlier blooming times.
+@sec-lit-review does a broad review of existing work surrounding cherry blossoms, their significance, and efforts to understand their behaviour related to climate.
+@sec-data covers the data sources, methodology, and conducts an exploratory analysis of the dataset.
+@sec-model proposes a model of cherry blossom flowering day as a function of mean March temperature in Kyoto and lists its results.
+@sec-discussion critically examines the data and results, comes to several conclusions about the dataset, and proposes new areas of exploration.
+This analysis highlights the need for...
+[FILL IN AT END]
 
-```
+# Literature Review {#sec-lit-review}
 
-This paper first focuses on quantifying a relationship between mean March temperature and the day of blooming of cherry blossom, using a mixture of historical data (@A)
-```
-
-# Literature Review
-
-```         
 The relationship between sakura blooming date and other factors was first examined in depth.
@@
 The study demonstrates that while cherry trees will continue to meet their chilling requirements under projected climate scenarios, accelerated heat accumulation will drive the dramatic advancement in flowering times.
-```
 
-# Research Question
+# Data {#sec-data}
 
-Clearly state your primary research question(s) and hypotheses if applicable.
+## Data Sources
 
-# Methodology
-
-## Research Design
-
-## Research Design
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Nullam eu justo in nisl tincidunt tincidunt.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Phasellus lacinia, urna eget venenatis vestibulum, nisi metus feugiat nisi, vel faucibus magna sem quis tellus.
-Donec eget ullamcorper magna, non lacinia erat.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Maecenas id eleifend enim.
-Cras rhoncus, magna in tempus tempor, dolor magna facilisis nisi, at ultrices velit mi quis orci.
-
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Vestibulum tortor quam, feugiat vitae, ultricies eget, tempor sit amet, ante.
-Donec eu libero sit amet quam egestas semper.
-Aenean ultricies mi vitae est.
-Mauris placerat eleifend leo.
-Quisque sit amet est et sapien ullamcorper pharetra.
-Vestibulum erat wisi, condimentum sed, commodo vitae, ornare sit amet, wisi.
-Aenean fermentum, elit eget tincidunt condimentum, eros ipsum rutrum orci, sagittis tempus lacus enim ac dui.
-Donec non enim in turpis pulvinar facilisis.
-Ut felis.
-Praesent dapibus, neque id cursus faucibus, tortor neque egestas augue, eu vulputate magna eros eu erat.
-Aliquam erat volutpat.
-Nam dui mi, tincidunt quis, accumsan porttitor, facilisis luctus, metus.
-Phasellus ultrices nulla quis nibh.
-Quisque a lectus.
-Donec consectetuer ligula vulputate sem tristique cursus.
-Nam nulla quam, gravida non, commodo a, sodales sit amet, nisi.
-Sed vitae justo eu ipsum maximus efficitur.
-Curabitur at dolor sed felis laoreet interdum vitae id mi.
-Fusce gravida pulvinar felis, a sagittis nisi sagittis non.
-
-## Analytical Approach
-
-# Data
-
-## Data Source, Collection, and Cleaning
+## Methodology
 
@@
 
-```{python}
-
-```
-
-# Modeling
+# Modeling {#sec-model}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-# Conclusion
+# References
 
-Summarize your key findings and their broader significance.
-
-# References
+# Appendix {#sec-appendix}

### 2025-03-31 paper/paper.qmd
wrote part of data section

@@
         geometry: "margin=1.2in" # print margins (change if wanting to)
-bibliography: references.bib
+bibliography: "references.bib"
 execute:
@@
 ---
+
+```{python}
+# Import necessary libraries 
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.dates as mdates
+import seaborn as sns
+import statsmodels.api as sm
+import numpy as np
+from matplotlib.colors import LinearSegmentedColormap
+from matplotlib.ticker import MaxNLocator
+from datetime import datetime
+
+# Load the historical data
+historic_bloom = pd.read_csv("../data/analysis_data/historic_bloom.csv")
+
+# Load the modern data
+modern_bloom = pd.read_csv("../data/analysis_data/modern_bloom.csv")
+
+# Truncate temperature to 1 decimal point
+historic_bloom['avg_temp_march_c'] = historic_bloom['avg_temp_march_c'].round(1)
+modern_bloom['avg_temp_march_c'] = modern_bloom['avg_temp_march_c'].round(1)
+
+```
 
@@
 
-In this paper, we analyse and model the nearly 1200-year record compiled by @Aono2010 along modern data compiled by the Japanese Meteorological Agency (JMA) [@JMA_sakura] [@JMA_temp], collected together by @Cookson.
-In particular, we seek to answer the question of whether warmer temperatures lead to earlier blooming times.
-@sec-lit-review does a broad review of existing work surrounding cherry blossoms, their significance, and efforts to understand their behaviour related to climate.
+We seek to answer the question of whether warmer temperatures lead to earlier blooming times.
+To do this, we analyse and model the nearly 1200-year record compiled by @Aono2010 in tandem with modern data compiled by the Japanese Meteorological Agency (JMA), collected and compiled together by @Cookson.@sec-lit-review does a broad review of existing work surrounding cherry blossoms, their significance, and efforts to understand their behaviour related to climate.
 @sec-data covers the data sources, methodology, and conducts an exploratory analysis of the dataset.
@@
 This analysis highlights the need for...
-[FILL IN AT END]
+\[FILL IN AT END\]
 
@@
 
+To investigate the relationship between mean March temperature and blooming date in Kyoto, historical data on cherry blossom blooming dates [@Aono2010], as well as data collected through modern measurement techniques [@JMA_sakura] [@JMA_temp] were analyzed.
+@Cookson provided both datasets assembled in a single repository, though the modern data sources were last updated in 2020.
+
+All of the data was transformed in Python [@python], using the `pandas` [@pandas] and `numpy` [@numpy] packages.
+For this analysis, the average (mean) march temperature, bloom date, day of year of blooming and year were made of interest.
+
+In the historical data, @Aono2010 reconstructs the mean march temperature in Kyoto for some of the years, based on the blooming date, but also provided observed temperatures taken from local newspapers for a given year when possible.
+We prioritized used observed temperatures, but if none were available, then the reconstructed one was used.
+This created a dataset of 827 observations, from the year 812-2015.
+A sample of the table can be seen in @tbl-historical.
+
+With the modern data, @Cookson combined average monthly temperatures and blooming date in a single dataset for several cities across Japan.
+We filtered this data to solely Kyoto, with average march temperatures to match the historical data.
+This created a dataset with 67 observations from the year 1953-2015.
+A sample can be seen in @tbl-modern.
+
+In our data exploration, the `matplotlib` [@matplotlib] and `seaborn` packages were used to visualize the data.
+
+```{python}
+#| label: tbl-historical
+#| tbl-cap: "Historical Cherry Blossom Blooming"
+
+# Rename columns for better readability
+historic_bloom_renamed = historic_bloom.rename(columns={
+    'year': 'Year',
+    'flower_date': 'Bloom Date',
+    'flower_doy': 'Day of Year',
+    'avg_temp_march_c': 'March Temp (°C)'
+})
+# Display the table without .style.set_caption()
+display(historic_bloom_renamed.head())
+```
+
+```{python}
+#| label: tbl-modern
+#| tbl-cap: "Modern Cherry Blossom Blooming"
+
+# Rename columns for better readability
+modern_bloom_renamed = modern_bloom.rename(columns={
+    'year': 'Year',
+    'flower_date': 'Bloom Date',
+    'flower_doy': 'Day of Year',
+    'avg_temp_march_c': 'March Temp (°C)'
+})
+
+# Display the table without .style.set_caption()
+display(modern_bloom_renamed.head())
+```
+
+The datasets were first explored chronologically, beginning with the historical data and ending with the modern data.
+
 ## Data Sources
@@
 
+asl;dkfj;laksdfjl;asjdf;lasjdfljasl;fj;lasjdf
+
 ## Model Specification
@@
 # Discussion {#sec-discussion}
+
+alsdhj;lasjfljas;lfja;lsjf;lasjdf;
 

### 2025-03-31 scripts/01-clean-data.py

@@ -17,12 +17,6 @@ df_modern_temp = pd.read_csv("data/raw_data/cookson_cleaned_csvs/temperatures-mo
 df_modern_blooming = pd.read_csv("data/raw_data/cookson_cleaned_csvs/sakura-modern.csv") # Load modern blooming data
 df_historical_temp = pd.read_csv("data/raw_data/cookson_cleaned_csvs/sakura-historical.csv") # Load historical temperature data
 
-print("Modern Temp data\n" + str(df_modern_temp.head())) # Print first 5 rows of modern temperature data
-print()
-print("Modern bloom data\n" + str(df_modern_blooming.head())) # Print first 5 rows of modern blooming data
-print()
-print("Historical temp/bloom data\n" + str(df_historical_temp.head())) # Print first 5 rows of historical temperature data
-
 ## TRANSFORM HISTORICAL DATA ##
 
 # Merge reconstructed temp with observed temp into one col (prioritize observed)
@@ -41,6 +35,9 @@ df_historical_temp["temp_c_recon"] = df_historical_temp["temp_c_recon"].astype("
 # drop unnecessary columns
 df_historical_temp = df_historical_temp.drop(columns=["study_source", "flower_source_name", "flower_source", "temp_c_obs", "temp_c_recon"])
 
+# remove years with no flowering data (NaN in flower_date)
+df_historical_temp = df_historical_temp.dropna(subset=["flower_date"])
+
 # reorder columns
 df_historical_temp = df_historical_temp[["year", "flower_date", "flower_doy", "temp_march"]]
 
@@ -49,6 +46,8 @@ df_historical_temp = df_historical_temp.rename(columns={"temp_march": "avg_temp_
 
 # save historical data to CSV file for one model
 df_historical_temp.to_csv("data/analysis_data/historic_bloom.csv", index=False) # Save historical data to CSV file
+
+
 
 
 ## TRANSFORM MODERN DATA ##
@@ -74,3 +73,4 @@ df_modern_bloom = df_modern_bloom.rename(columns={"mean_temp_c": "avg_temp_march
 ## save modern data to CSV file for one model
 df_modern_bloom.to_csv("data/analysis_data/modern_bloom.csv", index=False) # Save modern data to CSV file
 
+print(df_modern_bloom.shape)

### 2025-03-31 paper/paper.qmd
saved initial OLS model moved visualizations to paper

@@
 We seek to answer the question of whether warmer temperatures lead to earlier blooming times.
-To do this, we analyse and model the nearly 1200-year record compiled by @Aono2010 in tandem with modern data compiled by the Japanese Meteorological Agency (JMA), collected and compiled together by @Cookson.@sec-lit-review does a broad review of existing work surrounding cherry blossoms, their significance, and efforts to understand their behaviour related to climate.
+To do this, we analyse and model the nearly 1200-year record compiled by @Aono2010 in tandem with modern data compiled by the Japanese Meteorological Agency (JMA), collected and compiled together by @Cookson.
+@sec-lit-review does a broad review of existing work surrounding cherry blossoms, their significance, and efforts to understand their behaviour related to climate.
 @sec-data covers the data sources, methodology, and conducts an exploratory analysis of the dataset.
@@
 
+Boilerplate table
+
+```{python}
+
+```
+
+Boilerplate Fig
+
+```{python}
+
+```
+
 ```{python}
@@
 
-## Data Sources
+```{python}
 
-## Methodology
+historic = historic_bloom
+modern = modern_bloom
+
+# Drop NaN values
+historic = historic.dropna()
+modern = modern.dropna()
+
+```
+
+HISTORICAL DATA
+
+```{python}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: WRITE SOMETHING
+#| fig-align: center
+
+# Scatterplot of historic temp vs years 
+sns.set(style="whitegrid")
+plt.figure(figsize=(10, 6))
+
+# Create a custom pink/light blue color palette
+pink_blue_colors = ['#FF9ECD', '#87CEFA']  # Light pink and light blue
+custom_cmap = LinearSegmentedColormap.from_list('pink_blue', pink_blue_colors)
+
+# Scatter plot with pink color
+plt.scatter(historic['year'], historic['avg_temp_march_c'], alpha=0.7, color='#FF9ECD', edgecolor='#E75480', s=50)
+
+# Add moving average line in light blue
+historic['moving_avg'] = historic['avg_temp_march_c'].rolling(window=10).mean()
+plt.plot(historic['year'], historic['moving_avg'], color='#4BA3C3', linewidth=2.5, label='10-Year Moving Avg')
+
+# Add reference line for mean temperature
+mean_temp = historic['avg_temp_march_c'].mean()
+plt.axhline(y=mean_temp, color='#87CEFA', linestyle='--', alpha=0.8, 
+            label=f'Historical Mean ({mean_temp:.1f}°C)')
+
+plt.legend(framealpha=0.7)
+
+# Customize the plot with pink/blue theme
+plt.title('Historic Temperature vs Year', fontsize=16, color='#E75480')
+plt.xlabel('Year', fontsize=12, color='#4BA3C3')
+plt.ylabel('Temperature (°C)', fontsize=12, color='#4BA3C3')
+plt.xticks(rotation=45, color='#4F86C6')
+plt.yticks(color='#4F86C6')
+plt.grid(True, alpha=0.3, color='#E0E0E0')
+
+# Add pink/blue background touches
+plt.gca().set_facecolor('#F8F8FF')  # Light background
+plt.gcf().set_facecolor('#F8F8FF')
+
+plt.tight_layout()
+plt.show()
+```
+
+```{python}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: WRITE SOMETHING
+#| fig-align: center
+
+# Scatterplot of historic year vs flowering day
+plt.figure(figsize=(12, 7))
+
+# Filter out NaN values
+filtered_historic = historic.dropna(subset=['flower_doy'])
+
+# Create scatter plot with pink color
+plt.scatter(filtered_historic['year'], filtered_historic['flower_doy'], alpha=0.7, color='#FF9ECD', 
+            edgecolor='#E75480', s=50, label='Flowering Day')
+
+# Add moving average line with light blue color
+window_size = 10  # 10-year moving average
+filtered_historic['moving_avg_doy'] = filtered_historic['flower_doy'].rolling(window=window_size).mean()
+plt.plot(filtered_historic['year'], filtered_historic['moving_avg_doy'], color='#4BA3C3', linewidth=2.5, 
+         label=f'{window_size}-Year Moving Avg')
+
+# Add reference line for mean with light blue
+plt.axhline(y=mean_doy, color='#87CEFA', linestyle='--', alpha=0.8, 
+            label=f'Historical Mean ({mean_doy:.1f}, April 13-14)')
+
+# Customize the plot with pink/blue theme
+plt.title('Cherry Blossom Flowering Day Throughout History', fontsize=16, color='#E75480')
+plt.xlabel('Year', fontsize=12, color='#4BA3C3')
+plt.ylabel('Day of Year (DOY)', fontsize=12, color='#4BA3C3')
+plt.grid(True, alpha=0.3, color='#E0E0E0')
+plt.xticks(color='#4F86C6')
+plt.yticks(color='#4F86C6')
+plt.legend(framealpha=0.7)
+
+# Create a secondary y-axis with date labels
+ax = plt.gca()
+ax2 = ax.twinx()
+
+# Get the current y-limits from the primary axis
+y_min, y_max = ax.get_ylim()
+
+# Create properly aligned DOY-to-date mapping
+# In a non-leap year:
+# DOY 90 = March 31
+# DOY 91 = April 1
+# DOY 121 = May 1
+
+doy_values = np.arange(int(y_min), int(y_max) + 1, 5)  # Create ticks every 5 days
+date_labels = []
+
+for doy in doy_values:
+    if doy < 90:
+        # DOYs below 90 (March 31)
+        day_in_mar = doy - 59  # 59 = DOYs in Jan (31) + Feb (28)
+        date_labels.append(f'Mar {day_in_mar}')
+    elif doy < 120:
+        # DOYs for April (90-119)
+        day_in_apr = doy - 90 + 1
+        date_labels.append(f'Apr {day_in_apr}')
+    elif doy < 151:
+        # DOYs for May (120-150)
+        day_in_may = doy - 120 + 1
+        date_labels.append(f'May {day_in_may}')
+    else:
+        # DOYs for June and beyond
+        day_in_jun = doy - 151 + 1
+        date_labels.append(f'Jun {day_in_jun}')
+
+# Set the properly aligned ticks and labels
+ax2.set_ylim(y_min, y_max)
+ax2.set_yticks(doy_values)
+ax2.set_yticklabels(date_labels)
+ax2.set_ylabel('Approximate Date', color='#4BA3C3', fontsize=14)
+ax2.tick_params(axis='y', colors='#4F86C6')
+
+plt.legend(framealpha=0.7)
+plt.grid(True, alpha=0.3, axis='y')
+
+# Add pink/blue background touches
+plt.gca().set_facecolor('#F8F8FF')  # Light background
+plt.gcf().set_facecolor('#F8F8FF')
+
+plt.tight_layout()
+plt.show()
+```
+
+```{python}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: WRITE SOMETHING
+#| fig-align: center
+
+# Violin Plot
+
+# Create a figure
+plt.figure(figsize=(14, 8))
+
+# Define correct century order with 9th century first
+century_order = ['9th', '10th', '11th', '12th', '13th', '14th', '15th', 
+                '16th', '17th', '18th', '19th', '20th', '21st']
+
+# Create a custom color palette that transitions from light blue to pink
+start_color = '#ADD8E6'  # Light blue
+end_color = '#FFC0CB'    # Light pink
+
+# Generate a custom colormap
+n_colors = len(century_order)
+colors = [start_color, end_color]
+cmap_name = 'blue_to_pink'
+cm = LinearSegmentedColormap.from_list(cmap_name, colors, N=n_colors)
+
+# Get the colors for each century
+custom_palette = {}
+for i, century in enumerate(century_order):
+    custom_palette[century] = cm(i/float(n_colors-1))
+
+# Plot the violin plot with the custom order and palette
+sns.violinplot(x='century_label', y='flower_doy', data=historic, 
+               palette=custom_palette, inner='quartile', order=century_order)
+
+# Add individual data points with small jitter
+sns.stripplot(x='century_label', y='flower_doy', data=historic, 
+              color='#FF9ECD', alpha=0.3, jitter=True, size=4, order=century_order)
+
+# Add reference line for mean flowering day (assuming mean_doy is defined elsewhere)
+mean_doy = historic['flower_doy'].mean()  # Calculate if not already defined
+plt.axhline(y=mean_doy, color='#E75480', linestyle='--', alpha=0.8, 
+            label=f'Historical Mean ({mean_doy:.1f}, April 14)')
+
+# Enhance visuals
+plt.title('Cherry Blossom Flowering Day Distribution by Century', fontsize=18, color='#E75480')
+plt.xlabel('Century', fontsize=14, color='#4BA3C3')
+plt.ylabel('Day of Year (DOY)', fontsize=14, color='#4BA3C3')
+plt.xticks(rotation=45, color='#4F86C6')
+plt.yticks(color='#4F86C6')
+
+# Create a secondary y-axis with date labels
+ax = plt.gca()
+ax2 = ax.twinx()
+
+# Get the current y-limits from the primary axis
+y_min, y_max = ax.get_ylim()
+
+# Create properly aligned DOY-to-date mapping
+# In a non-leap year:
+# DOY 90 = March 31
+# DOY 91 = April 1
+# DOY 121 = May 1
+
+doy_values = np.arange(int(y_min), int(y_max) + 1, 5)  # Create ticks every 5 days
+date_labels = []
+
+for doy in doy_values:
+    if doy < 90:
+        # DOYs below 90 (March 31)
+        day_in_mar = doy - 59  # 59 = DOYs in Jan (31) + Feb (28)
+        date_labels.append(f'Mar {day_in_mar}')
+    elif doy < 120:
+        # DOYs for April (90-119)
+        day_in_apr = doy - 90 + 1
+        date_labels.append(f'Apr {day_in_apr}')
+    elif doy < 151:
+        # DOYs for May (120-150)
+        day_in_may = doy - 120 + 1
+        date_labels.append(f'May {day_in_may}')
+    else:
+        # DOYs for June and beyond
+        day_in_jun = doy - 151 + 1
+        date_labels.append(f'Jun {day_in_jun}')
+
+# Set the properly aligned ticks and labels
+ax2.set_ylim(y_min, y_max)
+ax2.set_yticks(doy_values)
+ax2.set_yticklabels(date_labels)
+ax2.set_ylabel('Approximate Date', color='#4BA3C3', fontsize=14)
+ax2.tick_params(axis='y', colors='#4F86C6')
+
+plt.legend(framealpha=0.7)
+plt.grid(True, alpha=0.3, axis='y')
+
+# Set background color
+plt.gca().set_facecolor('#F8F8FF')
+plt.gcf().set_facecolor('#F8F8FF')
+
+plt.tight_layout()
+plt.show()
+```
+
+MODERN DATA
+
+```{python}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: WRITE SOMETHING
+#| fig-align: center
+
+# Scatterplot of modern year vs flowering day
+plt.figure(figsize=(12, 7))
+
+# Filter out NaN values
+filtered_modern = modern.dropna(subset=['flower_doy'])
+
+# Create scatter plot with pink color
+plt.scatter(filtered_modern['year'], filtered_modern['flower_doy'], alpha=0.7, color='#FF9ECD', 
+            edgecolor='#E75480', s=50, label='Flowering Day')
+
+# Add moving average line with light blue color
+window_size = 5  # 5-year moving average
+filtered_modern['moving_avg_doy'] = filtered_modern['flower_doy'].rolling(window=window_size).mean()
+plt.plot(filtered_modern['year'], filtered_modern['moving_avg_doy'], color='#4BA3C3', linewidth=2.5, 
+         label=f'{window_size}-Year Moving Avg')
+
+# Customize the plot with pink/blue theme
+plt.title('Cherry Blossom Flowering Day Throughout Modern Times', fontsize=16, color='#E75480')
+plt.xlabel('Year', fontsize=12, color='#4BA3C3')
+plt.ylabel('Day of Year (DOY)', fontsize=12, color='#4BA3C3')
+plt.grid(True, alpha=0.3, color='#E0E0E0')
+plt.xticks(color='#4F86C6')
+plt.yticks(color='#4F86C6')
+plt.legend(framealpha=0.7)
+
+# Create a secondary y-axis with date labels
+ax = plt.gca()
+ax2 = ax.twinx()
+
+# Get the current y-limits from the primary axis
+y_min, y_max = ax.get_ylim()
+
+# Create properly aligned DOY-to-date mapping
+# In a non-leap year:
+# DOY 90 = March 31
+# DOY 91 = April 1
+# DOY 121 = May 1
+
+doy_values = np.arange(int(y_min), int(y_max) + 1, 5)  # Create ticks every 5 days
+date_labels = []
+
+for doy in doy_values:
+    if doy < 90:
+        # DOYs below 90 (March 31)
+        day_in_mar = doy - 59  # 59 = DOYs in Jan (31) + Feb (28)
+        date_labels.append(f'Mar {day_in_mar}')
+    elif doy < 120:
+        # DOYs for April (90-119)
+        day_in_apr = doy - 90 + 1
+        date_labels.append(f'Apr {day_in_apr}')
+    elif doy < 151:
+        # DOYs for May (120-150)
+        day_in_may = doy - 120 + 1
+        date_labels.append(f'May {day_in_may}')
+    else:
+        # DOYs for June and beyond
+        day_in_jun = doy - 151 + 1
+        date_labels.append(f'Jun {day_in_jun}')
+
+# Set the properly aligned ticks and labels
+ax2.set_ylim(y_min, y_max)
+ax2.set_yticks(doy_values)
+ax2.set_yticklabels(date_labels)
+ax2.set_ylabel('Approximate Date', color='#4BA3C3', fontsize=14)
+ax2.tick_params(axis='y', colors='#4F86C6')
+
+plt.legend(framealpha=0.7)
+plt.grid(True, alpha=0.3, axis='y')
+
+# Add pink/blue background touches
+plt.gca().set_facecolor('#F8F8FF')  # Light background
+plt.gcf().set_facecolor('#F8F8FF')
+
+plt.tight_layout()
+plt.show()
+```
+
+```{python}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: WRITE SOMETHING
+#| fig-align: center
+
+# scatterplot of modern data - year vs. Temperature
+
+# Scatterplot of modern temp vs years 
+sns.set(style="whitegrid")
+plt.figure(figsize=(10, 6))
+
+# Create a custom pink/light blue color palette
+pink_blue_colors = ['#FF9ECD', '#87CEFA']  # Light pink and light blue
+custom_cmap = LinearSegmentedColormap.from_list('pink_blue', pink_blue_colors)
+
+# Scatter plot with pink color
+plt.scatter(modern['year'], modern['avg_temp_march_c'], alpha=0.7, color='#FF9ECD', edgecolor='#E75480', s=50)
+
+# Add moving average line in light blue
+modern['moving_avg'] = modern['avg_temp_march_c'].rolling(window=5).mean()  # Using 5-year window for modern data
+plt.plot(modern['year'], modern['moving_avg'], color='#4BA3C3', linewidth=2.5, label='5-Year Moving Avg')
+
+plt.legend(framealpha=0.7)
+
+# Customize the plot with pink/blue theme
+plt.title('Modern Temperature vs Year', fontsize=16, color='#E75480')
+plt.xlabel('Year', fontsize=12, color='#4BA3C3')
+plt.ylabel('Temperature (°C)', fontsize=12, color='#4BA3C3')
+plt.xticks(rotation=45, color='#4F86C6')
+plt.yticks(color='#4F86C6')
+plt.grid(True, alpha=0.3, color='#E0E0E0')
+
+# Add pink/blue background touches
+plt.gca().set_facecolor('#F8F8FF')  # Light background
+plt.gcf().set_facecolor('#F8F8FF')
+
+plt.tight_layout()
+plt.show()
+```
+
+```{python}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: WRITE SOMETHING
+#| fig-align: center
+
+# histogram of flowering days
+plt.figure(figsize=(12, 7))
+plt.hist(modern['flower_doy'], bins=30, color='#FF9ECD', edgecolor='#E75480', alpha=0.7)
+plt.title('Distribution of Flowering Days (Modern)', fontsize=18, color='#E75480')
+plt.xlabel('Day of Year (DOY)', fontsize=14, color='#4BA3C3')
+plt.ylabel('Frequency', fontsize=14, color='#4BA3C3')
+plt.xticks(color='#4F86C6')
+plt.yticks(color='#4F86C6')
+plt.legend(framealpha=0.7)
+plt.grid(True, alpha=0.3, color='#E0E0E0')
+plt.gca().set_facecolor('#F8F8FF')  # Light background
+plt.gcf().set_facecolor('#F8F8FF')
+plt.tight_layout()
+plt.show()
+```
+
+```{python}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: WRITE SOMETHING
+#| fig-align: center
+
+# histogram of temperatures
+plt.figure(figsize=(12, 7))
+plt.hist(modern['avg_temp_march_c'], bins=30, color='#FF9ECD', edgecolor='#E75480', alpha=0.7)
+plt.title('Distribution of Average March Temperatures (Modern)', fontsize=18, color='#E75480')
+plt.xlabel('Temperature (°C)', fontsize=14, color='#4BA3C3')
+plt.ylabel('Frequency', fontsize=14, color='#4BA3C3')
+plt.xticks(color='#4F86C6')
+plt.yticks(color='#4F86C6')
+plt.legend(framealpha=0.7)
+plt.grid(True, alpha=0.3, color='#E0E0E0')
+plt.gca().set_facecolor('#F8F8FF')  # Light background
+plt.gcf().set_facecolor('#F8F8FF')
+plt.tight_layout()
+plt.show()
+
+```
 

### 2025-03-31 paper/paper.qmd
added descriptions for graphs

@@
 ```
+This scatterplot displays historical annual temperature records (in °C) from the year 800 CE to 2000 CE.
+Each pink dot represents the recorded temperature for a given year, while the solid teal line represents a 10-year moving average, smoothing out year-to-year fluctuations to highlight longer-term trends.
+The dashed horizontal line represents the historical mean temperature of 6.2°C.
+Notably, the data show considerable fluctuations over the centuries, with a relatively stable pattern through much of the pre-industrial era.
+However, a sharp and continuous rise in average temperature is observed beginning in the late 19th to early 20th century, suggesting significant warming recently.
+This trend aligns with broader patterns of climate change and may have implications for phenological events such as the timing of sakura (cherry blossom) flowering.
 
@@
 ```
+This graph showcases the annual full flowering dates of sakura from 800 to 2000 CE, plotted as day-of-year (DOY) values, where higher values indicate later blooming in the calendar year.
+Each pink dot corresponds to an individual year's observed full bloom date, and the solid teal line shows a 10-year moving average; the dashed horizontal line indicates the historical mean, which is 104.6.
+The data suggest that, while flowering dates have fluctuated across centuries, a notable trend toward earlier blooming began in the late 19th and 20th centuries.
+This shift coincides with the period of modern climate warming seen in the corresponding temperature graph, supporting the hypothesis that rising spring temperatures are advancing sakura bloom time.
 
@@
 ```
+The violin plot above showcases the cherry blossom flowering date distribution by century (9th to 21st), measured in day-of-year (DOY).
+The dashed pink line represents the average flowering day of the year (~105 days).
+It was quite consistent until the 20th century (1900s), where we can see a large drop; this trend has also been carried into the 21st century, where it even dropped lower.
 
@@
 ```
+The scatterplot above presents the annual full flowering dates of cherry blossoms in Japan from 1950 to 2020, measured in day-of-year (DOY), where higher values represent later blooming.
+Each pink dot marks the full flowering day for a given year, while the solid blue line shows a 5-year moving average, smoothing short-term fluctuations to reveal broader temporal trends.
+Over the 70-year span, there is a clear downward trend, indicating progressively earlier flowering.
 
@@
 ```
+This scatterplot shows the annual spring temperature in Japan from 1950 to 2020, measured in degrees Celsius (°C) relative to a baseline average.
+Each pink dot represents the average March temperature for a given year, and the solid blue line indicates a 5-year moving average to highlight underlying climate trends.
+The overall trajectory shows a clear increase in average March temperatures over the 70-year period, particularly after 1990, reflecting the impact of modern global climate change.
+This warming trend corresponds closely with the earlier flowering of cherry blossoms during the same time frame shown in the earlier figure.
+Together, both graphs reinforce the connection between rising spring temperatures and timing shifts in sakura blooming.
 
@@
 ```
+This histogram highlights the frequency distribution of sakura full flowering dates in modern Japan, measured in day-of-year (DOY), for the period between 1950 and 2020.
+Each bar represents how often flowering occurred within a specific DOY range.
+The distribution is roughly bell-shaped and unimodal, with the majority of flowering dates clustered between DOY 83 and DOY 95 (approximately March 24 to April 4), suggesting a central tendency in late March.
 
@@
 plt.show()
-
 ```
+The figure above presents the frequency distribution of average March temperatures in Japan from 1950 to 2020.
+The temperature values are plotted on the x-axis in degrees Celsius (°C).
+The distribution is centered around -1.5°C, with a wider spread toward colder extremes and a thinner tail extending into positive temperatures.
+While most March temperatures during this period fall below freezing, the presence of warmer years, particularly in the latter decades, indicates a shift in climate norms.
+These warming trends align with observed shifts in cherry blossom phenology, where earlier flowering is increasingly correlated with milder spring conditions.
 

### 2025-04-02 paper/paper.qmd
updated first draft of implication, future research direction, and limitation

@@
 ## Implications of Findings
+Our statistical analysis reveals a significant inverse relationship between pre-flowering March temperatures and sakura bloom dates, with higher temperatures consistently associated with earlier flowering time.
+This temperature-dependent pattern aligns with the documented global warming trend observed over recent centuries, particularly the accelerated temperature increase in the modern era.
+
+The predictive models generated through our research demonstrate robust statistical evidence that the progressive elevation of mean March temperatures, driven by global warming, will likely continue to advance cherry blossom flowering dates in coming decades.
+This phenological shift carries significant ecological, cultural, and economic implications.
+The alteration of these biorhythms may disrupt established ecological synchronizations between cherry trees and their pollinators, potentially affecting ecosystem functionality and biodiversity dynamics.
+
+From a cultural perspective, the temporal shift in cherry blossom seasons need adjustments to traditional hanami celebrations and tourism schedules, which have historically been calibrated to conventional blooming periods.
+In addition, these modifications could impact cultural practices that have remained relatively consistent for centuries and negatively impact tourism economy that is built around it.
 
 ## Limitations
+We would like to address several methodological limiting factors.
+Firstly, our analysis relies predominantly on data from Kyoto, representing a geographically limited subset that may not comprehensively reflect sakura phenological patterns across Japan's diverse climate regions.
+Secondly, although we used a merged dataset spanning from 800 CE to 2019, there are significant data gaps/missing values where we have to drop during the cleaning & merging of the two dataset.
+In addition, a notable limitation is the historical dataset's reliability, as large portions comprise of reconstructed values derived from local linear regression models or originated from primary sources that lack rigorous academic verification protocols.
+This presents challenges for establishing data provenance and reliability in accordance with contemporary scientific standards.
+Lastly, while our findings highlight temperature as a statistically significant predictor of bloom timing, the single-variable approach inherently overlooks other potential ecological factors.
+Variables such as precipitation patterns, photoperiod fluctuations, soil conditions, and winter chilling requirements may also contribute significantly to the observed advancement in flowering dates.
+A more comprehensive multivariate model would potentially yield enhanced predictive capacity and ecological insights.
 
@@
 ## Future Research Directions
+For future research, we would like to establish regression tree models to enhance predictive capabilities regarding sakura phenological patterns.
+Unlike traditional linear models, regression trees offer robust forecasting mechanisms that can better accommodate complex interactions between variables, prioritizing predictive accuracy over interpretive simplicity.
+Next, we recommend adding in additional environmental parameters beyond temperature.
+Variables warranting investigation including but not limited to precipitation frequency and intensity, atmospheric humidity gradients, soil composition metrics, photoperiod fluctuations, and cumulative chilling units during dormancy periods.
+Such multidimensional modeling would potentially explain a greater proportion of the observed variability in bloom timing and provide more nuanced insights into the complex biological mechanisms governing cherry blossom phenology.
+Lastly, future research could benefit from expanding the geographical scope beyond Kyoto to establish comparative phenological datasets across diverse microclimates and latitudinal gradients throughout Japan and other regions where cherry trees are important.
+This expanded spatial analysis would facilitate the identification of region-specific responses to climate variables and potentially isolate genetic adaptations to local environmental conditions.
 

### 2025-04-02 paper/paper.qmd
fixed modeling amended intro with implications summary

--- 2025-03-31 paper/paper.qmd
@@
 ---
-title: "Your Paper Title"
+title: "Warming Degrees and Flowering Trees"
 subtitle: "Understanding Temperature's Role in Earlier Cherry Blossom Blooming Times in Kyoto, Japan"
@@
 ```{python}
+#| label: data-setup
+
 # Import necessary libraries 
@@
 
+historic = historic_bloom
+modern = modern_bloom
+
+# Drop NaN values
+historic = historic.dropna()
+modern = modern.dropna()
+
 ```
@@
 @sec-discussion critically examines the data and results, comes to several conclusions about the dataset, and proposes new areas of exploration.
-This analysis highlights the need for...
-\[FILL IN AT END\]
+This analysis highlights several conclusions.
+First, temperature changes will have consequential effects for wildlife that function around the timing of blooming *sakura*.
+Next, social celebrations such as *hanami* will be impacted, changing the timing of a once-constant annual event.
+Finally, there is a need for more a more comprehensive inquiry into the determinants of cherry blossom blooming in Japan more generally, involving more covariates, so that more robust predictions of *sakura* blooming in Kyoto can be made.
+By gathering more data to make better predictions, the city of Kyoto will better be able to weather climate change and bridge their cherished traditions into the future.
 
@@
 
-With the modern data, @Cookson combined average monthly temperatures and blooming date in a single dataset for several cities across Japan.
-We filtered this data to solely Kyoto, with average march temperatures to match the historical data.
-This created a dataset with 67 observations from the year 1953-2015.
-A sample can be seen in @tbl-modern.
-
-In our data exploration, the `matplotlib` [@matplotlib] and `seaborn` packages were used to visualize the data.
-
-Boilerplate table
-
 ```{python}
-
-```
-
-Boilerplate Fig
-
-```{python}
-
-```
-
-```{python}
-#| label: tbl-historical
+#| label: tbl-historical-data
 #| tbl-cap: "Historical Cherry Blossom Blooming"
@@
 
+With the modern data, @Cookson combined average monthly temperatures and blooming date in a single dataset for several cities across Japan.
+We filtered this data to solely Kyoto, and solely average march temperatures to match the historical data from @Aono2010.
+This created a dataset with 67 observations ranging from 1953 to 2015.
+A sample can be seen in @tbl-modern.
+
 ```{python}
-#| label: tbl-modern
+#| label: tbl-modern-data
 #| tbl-cap: "Modern Cherry Blossom Blooming"
@@
 
+In our data exploration, the `matplotlib` [@matplotlib] and `seaborn` packages were used to visualize and explore the data.
 The datasets were first explored chronologically, beginning with the historical data and ending with the modern data.
-
-```{python}
-
-historic = historic_bloom
-modern = modern_bloom
-
-# Drop NaN values
-historic = historic.dropna()
-modern = modern.dropna()
-
-```
 
@@
 ```{python}
+#| label: historic-temp-yr-scatter
 #| message: false
@@
 
-## Implications of Findings
+## Ecological Implications
+
+## Cultural Implications
 
 ## Limitations
-
-## Robustness Checks
 

### 2025-04-02 paper/paper.qmd

--- 2025-04-02 paper/paper.qmd
@@
 
-alsdhj;lasjfljas;lfja;lsjf;lasjdf;
-
 ## Implications of Findings
@@
 
-## Robustness Checks
+## Ecological Implications
+
+## Cultural Implications
+
+## Limitations
 

### 2025-04-03 paper/paper.qmd
fixed diagrams in eda

@@
 The Japanese tradition of *hanami* (loosely translated as 'flower-viewing') is an important cultural tradition that honors this perspective, being associated with the flowering birth of cherry blossom trees in the late spring.
-In the city of Kyoto in particular, local newspapers have recorded the dates of tree flowering since the late 19th century, but researchers like @Aono2010 have been able to estimate flowering dates as far back as the year 812 CE, by combing through historical records such as diaries, poetry and other written documents describing *hanami* parties throughout the ages.
-However, rising global temperatures create an anomaly, highlighting the need to understand the historical and modern links between temperature and cherry blossom bloom timing.
-Without it, cherished traditions like *hanami* may struggle to adapt to climate shifts, risking their disappearance as a unique Japanese custom.
-To this end, work like @Allen2014's analyse and relatively modern blooming data over a 29-year period, but the key weakness of this is a lack of perspective.
-Given that their data only covers a small part of the entire recorded history of cherry blossom blooming, their analysis and subsequent model have a key weakness - they do not take into account large scale trends of blooming and their relationship to the change in climate that has happened in the last 150 years, which @Aono2010's data can illustrate.
+In the city of Kyoto in particular, local newspapers have recorded the dates of tree flowering since the late 19th century, but researchers like @Aono2010 have been able to corroborate flowering dates as far back as the year 812 CE, through the examination of historical records such as diaries, poetry and other written documents describing *hanami* parties throughout the ages.
+However, rising global temperatures have created anomalous climate conditions in recent years.
+This highlights a need to understand the historical and modern links between temperature and cherry blossom bloom timing.
+Without it, cherished traditions like *hanami* may struggle to adapt to climate shifts, risking their disappearance as unique customs.
 
-We seek to answer the question of whether warmer temperatures lead to earlier blooming times.
-To do this, we analyse and model the nearly 1200-year record compiled by @Aono2010 in tandem with modern data compiled by the Japanese Meteorological Agency (JMA), collected and compiled together by @Cookson.
+Work such as @Allen2014 analyzes cherry blossom blooming data over a relatively brief 29-year timespan.
+However, this narrow temporal window represents only a fraction of the documented flowering history, limiting their model's ability to capture the relationship between blossoms and the significant climate shifts of the past 150 years - shifts that are clearly visible in @Aono2010's more comprehensive dataset.
+
+We also seek to answer the question of whether warmer temperatures lead to earlier blooming times.
+To achieve this, we instead analyse and model the nearly 1200-year record compiled by @Aono2010 in tandem with modern data compiled by the Japanese Meteorological Agency (JMA), collected and compiled together by @Cookson.
 @sec-lit-review does a broad review of existing work surrounding cherry blossoms, their significance, and efforts to understand their behaviour related to climate.
@@
 @sec-discussion critically examines the data and results, comes to several conclusions about the dataset, and proposes new areas of exploration.
-This analysis highlights several conclusions.
-First, temperature changes will have consequential effects for wildlife that function around the timing of blooming *sakura*.
-Next, social celebrations such as *hanami* will be impacted, changing the timing of a once-constant annual event.
-Finally, there is a need for more a more comprehensive inquiry into the determinants of cherry blossom blooming in Japan more generally, involving more covariates, so that more robust predictions of *sakura* blooming in Kyoto can be made.
-By gathering more data to make better predictions, the city of Kyoto will better be able to weather climate change and bridge their cherished traditions into the future.
+This analysis highlights the  consequential effects for wildlife that function around the timing of blooming *sakura*, the need for flexibility in planning *hanami* celebrations going forward, as well as the need for wider ranging data to inform more precise predictions of blooming days.
 
@@
 ```{python}
-#| label: tbl-historical-data
+#| label: tbl-historical
 #| tbl-cap: "Historical Cherry Blossom Blooming"
@@
 })
+
 # Display the table without .style.set_caption()
@@
 ```{python}
-#| label: tbl-modern-data
+#| label: tbl-modern
 #| tbl-cap: "Modern Cherry Blossom Blooming"
@@
 
-In our data exploration, the `matplotlib` [@matplotlib] and `seaborn` packages were used to visualize and explore the data.
-The datasets were first explored chronologically, beginning with the historical data and ending with the modern data.
-
-HISTORICAL DATA
+In our data exploration, the `matplotlib` [@matplotlib] and `seaborn` [@seaborn] packages were used to visualize and explore the data. The historical dataset was visualized and explored first.
 
 ```{python}
-#| label: historic-temp-yr-scatter
+#| label: fig-hist-temp-yr-scatter
 #| message: false
@@
 #| warning: false
-#| fig-cap: WRITE SOMETHING
 #| fig-align: center
@@
 # Scatterplot of historic temp vs years 
-sns.set(style="whitegrid")
 plt.figure(figsize=(10, 6))
 
-# Create a custom pink/light blue color palette
-pink_blue_colors = ['#FF9ECD', '#87CEFA']  # Light pink and light blue
-custom_cmap = LinearSegmentedColormap.from_list('pink_blue', pink_blue_colors)
-
-# Scatter plot with pink color
+# Scatter plot with pink points
 plt.scatter(historic['year'], historic['avg_temp_march_c'], alpha=0.7, color='#FF9ECD', edgecolor='#E75480', s=50)
 
-# Add moving average line in light blue
-historic['moving_avg'] = historic['avg_temp_march_c'].rolling(window=10).mean()
-plt.plot(historic['year'], historic['moving_avg'], color='#4BA3C3', linewidth=2.5, label='10-Year Moving Avg')
+# Add moving average line
+moving_avg = historic['avg_temp_march_c'].rolling(window=10).mean()
+plt.plot(historic['year'], moving_avg, color='#5F3E2A', linewidth=2.5, label='10-Year Moving Avg')
 
@@
 mean_temp = historic['avg_temp_march_c'].mean()
-plt.axhline(y=mean_temp, color='#87CEFA', linestyle='--', alpha=0.8, 
-            label=f'Historical Mean ({mean_temp:.1f}°C)')
+plt.axhline(y=mean_temp, color='skyblue', linestyle='--', label=f'Historical Mean ({mean_temp:.1f}°C)')
 
-plt.legend(framealpha=0.7)
-
-# Customize the plot with pink/blue theme
-plt.title('Historic Temperature vs Year', fontsize=16, color='#E75480')
-plt.xlabel('Year', fontsize=12, color='#4BA3C3')
-plt.ylabel('Temperature (°C)', fontsize=12, color='#4BA3C3')
-plt.xticks(rotation=45, color='#4F86C6')
-plt.yticks(color='#4F86C6')
-plt.grid(True, alpha=0.3, color='#E0E0E0')
-
-# Add pink/blue background touches
-plt.gca().set_facecolor('#F8F8FF')  # Light background
-plt.gcf().set_facecolor('#F8F8FF')
+# Add labels and customize
+plt.title('Historic Temperature vs Year', fontsize=16)
+plt.xlabel('Year', fontsize=12)
+plt.ylabel('Temperature (°C)', fontsize=12)
+plt.grid(True, alpha=0.3)
+plt.legend()
 
 plt.tight_layout()
-plt.show()
 ```
-This scatterplot displays historical annual temperature records (in °C) from the year 800 CE to 2000 CE.
+
+@fig-hist-temp-yr-scatter displays historical annual temperature records (in °C) from the year 800 CE to 2000 CE.
 Each pink dot represents the recorded temperature for a given year, while the solid teal line represents a 10-year moving average, smoothing out year-to-year fluctuations to highlight longer-term trends.
@@
 ```{python}
+#| label: fig-hist-bloomday-yr
 #| message: false
@@
 #| warning: false
-#| fig-cap: WRITE SOMETHING
 #| fig-align: center
 
-# Scatterplot of historic year vs flowering day
-plt.figure(figsize=(12, 7))
+# Create simplified flowering day history plot
+plt.figure(figsize=(10, 6))
 
@@
 
-# Create scatter plot with pink color
-plt.scatter(filtered_historic['year'], filtered_historic['flower_doy'], alpha=0.7, color='#FF9ECD', 
-            edgecolor='#E75480', s=50, label='Flowering Day')
+# Create scatter plot 
+plt.scatter(filtered_historic['year'], filtered_historic['flower_doy'], 
+            alpha=0.7, color='#FF9ECD', edgecolor='#E75480', s=40)
 
-# Add moving average line with light blue color
-window_size = 10  # 10-year moving average
-filtered_historic['moving_avg_doy'] = filtered_historic['flower_doy'].rolling(window=window_size).mean()
-plt.plot(filtered_historic['year'], filtered_historic['moving_avg_doy'], color='#4BA3C3', linewidth=2.5, 
-         label=f'{window_size}-Year Moving Avg')
+# Add moving average line
+moving_avg = filtered_historic['flower_doy'].rolling(window=10).mean()
+plt.plot(filtered_historic['year'], moving_avg, 
+         color='#6E4B34', linewidth=2, label='10-Year Moving Avg')
 
-# Add reference line for mean with light blue
-plt.axhline(y=mean_doy, color='#87CEFA', linestyle='--', alpha=0.8, 
+# Add reference line for historical mean
+mean_doy = filtered_historic['flower_doy'].mean()
+plt.axhline(y=mean_doy, color='#87CEFA', linestyle='--', 
             label=f'Historical Mean ({mean_doy:.1f}, April 13-14)')
 
-# Customize the plot with pink/blue theme
-plt.title('Cherry Blossom Flowering Day Throughout History', fontsize=16, color='#E75480')
-plt.xlabel('Year', fontsize=12, color='#4BA3C3')
-plt.ylabel('Day of Year (DOY)', fontsize=12, color='#4BA3C3')
-plt.grid(True, alpha=0.3, color='#E0E0E0')
-plt.xticks(color='#4F86C6')
-plt.yticks(color='#4F86C6')
-plt.legend(framealpha=0.7)
-
-# Create a secondary y-axis with date labels
+# Set up secondary y-axis for date labels
 ax = plt.gca()
 ax2 = ax.twinx()
+y_min, y_max = ax.get_ylim()
+y_min, y_max = int(y_min), int(y_max)
 
-# Get the current y-limits from the primary axis
-y_min, y_max = ax.get_ylim()
-
-# Create properly aligned DOY-to-date mapping
-# In a non-leap year:
-# DOY 90 = March 31
-# DOY 91 = April 1
-# DOY 121 = May 1
-
-doy_values = np.arange(int(y_min), int(y_max) + 1, 5)  # Create ticks every 5 days
+# Create date labels
+doy_values = np.arange(y_min - (y_min % 10), y_max + 10, 10)  # Ticks every 10 days
 date_labels = []
+months = {3: 'Mar', 4: 'Apr', 5: 'May', 6: 'Jun'}
 
 for doy in doy_values:
+    # Convert DOY to month/day
+    if doy <= 90:
+        date_labels.append(f'{months[3]} {doy-59}')
+    elif doy <= 120:
+        date_labels.append(f'{months[4]} {doy-90+1}')
+    elif doy <= 151:
+        date_labels.append(f'{months[5]} {doy-120+1}')
+    else:
+        date_labels.append(f'{months[6]} {doy-151+1}')
+
+# Apply secondary axis settings
+ax2.set_ylim(y_min, y_max)
+ax2.set_yticks(doy_values)
+ax2.set_yticklabels(date_labels)
+ax2.set_ylabel('Approximate Date', color='#4BA3C3')
+
+# Customize appearance
+plt.title('Cherry Blossom Flowering Day Throughout History', fontsize=14)
+plt.xlabel('Year')
+plt.ylabel('Day of Year (DOY)')
+plt.grid(True, alpha=0.3)
+plt.legend()
+plt.tight_layout()
+```
+
+@fig-hist-bloomday-yr showcases the annual full flowering dates of sakura from 800 to 2000 CE, plotted as day-of-year (DOY) values, where higher values indicate later blooming in the calendar year.
+Each pink dot corresponds to an individual year's observed full bloom date, and the solid brown line shows a 10-year moving average; the dashed horizontal line indicates the historical mean, which is 104.6.
+The data suggest that while flowering dates have fluctuated across centuries, a notable trend toward earlier blooming began in the late 19th and 20th centuries.
+This shift coincides with the period of modern climate warming seen in the corresponding temperature graph, suggesting that rising spring temperatures are advancing sakura bloom time.
+
+```{python}
+#| label: fig-hist-violin-plot  
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Violin Plot 
+plt.figure(figsize=(14, 8))
+
+# Create century labels based on year
+if 'century_label' not in historic.columns:
+    century_num = (historic['year'] // 100 + 1).astype(int)
+    historic['century_label'] = century_num.apply(lambda x: f"{x}{'th' if x not in [1, 2, 3, 21] else ['st', 'nd', 'rd', 'st'][x % 10 - 1 if x % 100 != 11 else 0]}")
+
+# Define centuries in chronological order
+century_order = [f"{i}{'th' if i not in [21] else 'st'}" for i in range(9, 22)]
+
+# Create a color gradient from light blue to pink
+colors = ['#ADD8E6', '#FFC0CB']  # Light blue to Light pink
+cm = LinearSegmentedColormap.from_list('blue_to_pink', colors, N=len(century_order))
+custom_palette = {century: cm(i/float(len(century_order)-1)) for i, century in enumerate(century_order)}
+
+# Create violin plot
+sns.violinplot(x='century_label', y='flower_doy', data=historic, 
+               palette=custom_palette, inner='quartile', order=century_order)
+
+# Add data points
+sns.stripplot(x='century_label', y='flower_doy', data=historic, 
+              color='#FF9ECD', alpha=0.3, jitter=True, size=4, order=century_order)
+
+# Add mean reference line
+mean_doy = historic['flower_doy'].mean()
+plt.axhline(y=mean_doy, color='#E75480', linestyle='--', 
+            label=f'Historical Mean ({mean_doy:.1f}, April 14)')
+
+# Set titles and labels
+plt.title('Cherry Blossom Flowering Day Distribution by Century', fontsize=18, color='#E75480')
+plt.xlabel('Century', fontsize=14, color='#4BA3C3')
+plt.ylabel('Day of Year (DOY)', fontsize=14, color='#4BA3C3')
+plt.xticks(rotation=45, color='#4F86C6')
+
+# Add secondary y-axis with calendar dates
+ax = plt.gca()
+ax2 = ax.twinx()
+y_min, y_max = ax.get_ylim()
+doy_values = range(int(y_min), int(y_max) + 1, 10)  # Ticks every 10 days
+
+# Create date labels
+date_labels = []
+for doy in doy_values:
     if doy < 90:
-        # DOYs below 90 (March 31)
-        day_in_mar = doy - 59  # 59 = DOYs in Jan (31) + Feb (28)
-        date_labels.append(f'Mar {day_in_mar}')
+        date_labels.append(f'Mar {doy - 59}')
     elif doy < 120:
-        # DOYs for April (90-119)
-        day_in_apr = doy - 90 + 1
-        date_labels.append(f'Apr {day_in_apr}')
+        date_labels.append(f'Apr {doy - 90 + 1}')
     elif doy < 151:
-        # DOYs for May (120-150)
-        day_in_may = doy - 120 + 1
-        date_labels.append(f'May {day_in_may}')
+        date_labels.append(f'May {doy - 120 + 1}')
     else:
-        # DOYs for June and beyond
-        day_in_jun = doy - 151 + 1
-        date_labels.append(f'Jun {day_in_jun}')
+        date_labels.append(f'Jun {doy - 151 + 1}')
 
-# Set the properly aligned ticks and labels
 ax2.set_ylim(y_min, y_max)
@@
 ax2.set_ylabel('Approximate Date', color='#4BA3C3', fontsize=14)
-ax2.tick_params(axis='y', colors='#4F86C6')
 
-plt.legend(framealpha=0.7)
+plt.legend()
 plt.grid(True, alpha=0.3, axis='y')
+plt.tight_layout()
+```
 
-# Add pink/blue background touches
-plt.gca().set_facecolor('#F8F8FF')  # Light background
-plt.gcf().set_facecolor('#F8F8FF')
+@fig-hist-violin-plot showcases the cherry blossom flowering date distribution by century (9th to 21st), measured in day-of-year (DOY).
+The dashed pink line represents the average flowering day of the year (~105 days).
+It was quite consistent until the 20th century (1900s), where we can see a large drop; this trend has also been carried into the 21st century, where it dropped further.
 
-plt.tight_layout()
-plt.show()
-```
-This graph showcases the annual full flowering dates of sakura from 800 to 2000 CE, plotted as day-of-year (DOY) values, where higher values indicate later blooming in the calendar year.
-Each pink dot corresponds to an individual year's observed full bloom date, and the solid teal line shows a 10-year moving average; the dashed horizontal line indicates the historical mean, which is 104.6.
-The data suggest that, while flowering dates have fluctuated across centuries, a notable trend toward earlier blooming began in the late 19th and 20th centuries.
-This shift coincides with the period of modern climate warming seen in the corresponding temperature graph, supporting the hypothesis that rising spring temperatures are advancing sakura bloom time.
+A common thread emerges across all of the plots - warmer temperatures as time goes on has led to cherry blossoms blooming earlier and earlier every year.Furthermorethe significant shifting blooming days relative to temperature point to this historical data being ineffective in understanding the present relationship between temperature and blooming day.
+Since blooming days and temperatures have shifted so much in the last 150 years, examining more modern data is more effective in uncovering current trends.
+Given this, another more current dataset from @Cookson was examined next.
 
 ```{python}
+#| label: fig-modern-year-temp 
 #| message: false
@@
 #| warning: false
-#| fig-cap: WRITE SOMETHING
-#| fig-align: center
-
-# Violin Plot
-
-# Create a figure
-plt.figure(figsize=(14, 8))
-
-# Define correct century order with 9th century first
-century_order = ['9th', '10th', '11th', '12th', '13th', '14th', '15th', 
-                '16th', '17th', '18th', '19th', '20th', '21st']
-
-# Create a custom color palette that transitions from light blue to pink
-start_color = '#ADD8E6'  # Light blue
-end_color = '#FFC0CB'    # Light pink
-
-# Generate a custom colormap
-n_colors = len(century_order)
-colors = [start_color, end_color]
-cmap_name = 'blue_to_pink'
-cm = LinearSegmentedColormap.from_list(cmap_name, colors, N=n_colors)
-
-# Get the colors for each century
-custom_palette = {}
-for i, century in enumerate(century_order):
-    custom_palette[century] = cm(i/float(n_colors-1))
-
-# Plot the violin plot with the custom order and palette
-sns.violinplot(x='century_label', y='flower_doy', data=historic, 
-               palette=custom_palette, inner='quartile', order=century_order)
-
-# Add individual data points with small jitter
-sns.stripplot(x='century_label', y='flower_doy', data=historic, 
-              color='#FF9ECD', alpha=0.3, jitter=True, size=4, order=century_order)
-
-# Add reference line for mean flowering day (assuming mean_doy is defined elsewhere)
-mean_doy = historic['flower_doy'].mean()  # Calculate if not already defined
-plt.axhline(y=mean_doy, color='#E75480', linestyle='--', alpha=0.8, 
-            label=f'Historical Mean ({mean_doy:.1f}, April 14)')
-
-# Enhance visuals
-plt.title('Cherry Blossom Flowering Day Distribution by Century', fontsize=18, color='#E75480')
-plt.xlabel('Century', fontsize=14, color='#4BA3C3')
-plt.ylabel('Day of Year (DOY)', fontsize=14, color='#4BA3C3')
-plt.xticks(rotation=45, color='#4F86C6')
-plt.yticks(color='#4F86C6')
-
-# Create a secondary y-axis with date labels
-ax = plt.gca()
-ax2 = ax.twinx()
-
-# Get the current y-limits from the primary axis
-y_min, y_max = ax.get_ylim()
-
-# Create properly aligned DOY-to-date mapping
-# In a non-leap year:
-# DOY 90 = March 31
-# DOY 91 = April 1
-# DOY 121 = May 1
-
-doy_values = np.arange(int(y_min), int(y_max) + 1, 5)  # Create ticks every 5 days
-date_labels = []
-
-for doy in doy_values:
-    if doy < 90:
-        # DOYs below 90 (March 31)
-        day_in_mar = doy - 59  # 59 = DOYs in Jan (31) + Feb (28)
-        date_labels.append(f'Mar {day_in_mar}')
-    elif doy < 120:
-        # DOYs for April (90-119)
-        day_in_apr = doy - 90 + 1
-        date_labels.append(f'Apr {day_in_apr}')
-    elif doy < 151:
-        # DOYs for May (120-150)
-        day_in_may = doy - 120 + 1
-        date_labels.append(f'May {day_in_may}')
-    else:
-        # DOYs for June and beyond
-        day_in_jun = doy - 151 + 1
-        date_labels.append(f'Jun {day_in_jun}')
-
-# Set the properly aligned ticks and labels
-ax2.set_ylim(y_min, y_max)
-ax2.set_yticks(doy_values)
-ax2.set_yticklabels(date_labels)
-ax2.set_ylabel('Approximate Date', color='#4BA3C3', fontsize=14)
-ax2.tick_params(axis='y', colors='#4F86C6')
-
-plt.legend(framealpha=0.7)
-plt.grid(True, alpha=0.3, axis='y')
-
-# Set background color
-plt.gca().set_facecolor('#F8F8FF')
-plt.gcf().set_facecolor('#F8F8FF')
-
-plt.tight_layout()
-plt.show()
-```
-The violin plot above showcases the cherry blossom flowering date distribution by century (9th to 21st), measured in day-of-year (DOY).
-The dashed pink line represents the average flowering day of the year (~105 days).
-It was quite consistent until the 20th century (1900s), where we can see a large drop; this trend has also been carried into the 21st century, where it even dropped lower.
-
-MODERN DATA
-
-```{python}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: WRITE SOMETHING
 #| fig-align: center
@@
 
-# Filter out NaN values
-filtered_modern = modern.dropna(subset=['flower_doy'])
+# Create scatter plot
+plt.scatter(modern['year'], modern['flower_doy'], 
+            alpha=0.7, color='#FF9ECD', edgecolor='#E75480', s=50)
 
-# Create scatter plot with pink color
-plt.scatter(filtered_modern['year'], filtered_modern['flower_doy'], alpha=0.7, color='#FF9ECD', 
-            edgecolor='#E75480', s=50, label='Flowering Day')
+# Add moving average line
+moving_avg = modern['flower_doy'].rolling(window=5).mean()
+plt.plot(modern['year'], moving_avg, color='#6E4B34', linewidth=2.5, 
+         label='5-Year Moving Avg')
 
-# Add moving average line with light blue color
-window_size = 5  # 5-year moving average
-filtered_modern['moving_avg_doy'] = filtered_modern['flower_doy'].rolling(window=window_size).mean()
-plt.plot(filtered_modern['year'], filtered_modern['moving_avg_doy'], color='#4BA3C3', linewidth=2.5, 
-         label=f'{window_size}-Year Moving Avg')
-
-# Customize the plot with pink/blue theme
+# Customize plot appearance
 plt.title('Cherry Blossom Flowering Day Throughout Modern Times', fontsize=16, color='#E75480')
@@
 plt.ylabel('Day of Year (DOY)', fontsize=12, color='#4BA3C3')
-plt.grid(True, alpha=0.3, color='#E0E0E0')
-plt.xticks(color='#4F86C6')
-plt.yticks(color='#4F86C6')
-plt.legend(framealpha=0.7)
+plt.grid(True, alpha=0.3)
 
-# Create a secondary y-axis with date labels
+# Create secondary y-axis with date labels
 ax = plt.gca()
 ax2 = ax.twinx()
-
-# Get the current y-limits from the primary axis
 y_min, y_max = ax.get_ylim()
-
-# Create properly aligned DOY-to-date mapping
-# In a non-leap year:
-# DOY 90 = March 31
-# DOY 91 = April 1
-# DOY 121 = May 1
-
-doy_values = np.arange(int(y_min), int(y_max) + 1, 5)  # Create ticks every 5 days
+doy_values = np.arange(int(y_min), int(y_max) + 1, 5)
 date_labels = []
@@
     if doy < 90:
-        # DOYs below 90 (March 31)
-        day_in_mar = doy - 59  # 59 = DOYs in Jan (31) + Feb (28)
-        date_labels.append(f'Mar {day_in_mar}')
+        date_labels.append(f'Mar {doy-59}')
     elif doy < 120:
-        # DOYs for April (90-119)
-        day_in_apr = doy - 90 + 1
-        date_labels.append(f'Apr {day_in_apr}')
+        date_labels.append(f'Apr {doy-90+1}')
     elif doy < 151:
-        # DOYs for May (120-150)
-        day_in_may = doy - 120 + 1
-        date_labels.append(f'May {day_in_may}')
+        date_labels.append(f'May {doy-120+1}')
     else:
-        # DOYs for June and beyond
-        day_in_jun = doy - 151 + 1
-        date_labels.append(f'Jun {day_in_jun}')
+        date_labels.append(f'Jun {doy-151+1}')
 
-# Set the properly aligned ticks and labels
 ax2.set_ylim(y_min, y_max)
@@
 ax2.set_yticklabels(date_labels)
-ax2.set_ylabel('Approximate Date', color='#4BA3C3', fontsize=14)
-ax2.tick_params(axis='y', colors='#4F86C6')
+ax2.set_ylabel('Approximate Date', color='#4BA3C3')
 
-plt.legend(framealpha=0.7)
-plt.grid(True, alpha=0.3, axis='y')
+plt.legend()
+plt.tight_layout()
+```
 
-# Add pink/blue background touches
-plt.gca().set_facecolor('#F8F8FF')  # Light background
-plt.gcf().set_facecolor('#F8F8FF')
-
-plt.tight_layout()
-plt.show()
-```
-The scatterplot above presents the annual full flowering dates of cherry blossoms in Japan from 1950 to 2020, measured in day-of-year (DOY), where higher values represent later blooming.
+@fig-modern-year-temp above presents the annual full flowering dates of cherry blossoms in Japan from 1950 to 2020, measured in day-of-year (DOY), where higher values represent later blooming.
 Each pink dot marks the full flowering day for a given year, while the solid blue line shows a 5-year moving average, smoothing short-term fluctuations to reveal broader temporal trends.
@@
 ```{python}
+#| label: fig-modern-bloomday-yr
 #| message: false
@@
 modern['moving_avg'] = modern['avg_temp_march_c'].rolling(window=5).mean()  # Using 5-year window for modern data
-plt.plot(modern['year'], modern['moving_avg'], color='#4BA3C3', linewidth=2.5, label='5-Year Moving Avg')
+plt.plot(modern['year'], modern['moving_avg'], color='#6E4B34', linewidth=2.5, label='5-Year Moving Avg')
 
@@
 
-# Add pink/blue background touches
-plt.gca().set_facecolor('#F8F8FF')  # Light background
-plt.gcf().set_facecolor('#F8F8FF')
+plt.tight_layout()
+```
 
-plt.tight_layout()
-plt.show()
-```
-This scatterplot shows the annual spring temperature in Japan from 1950 to 2020, measured in degrees Celsius (°C) relative to a baseline average.
+@fig-modern-bloomday-yr shows the annual spring temperature in Japan from 1950 to 2020, measured in degrees Celsius (°C) relative to a baseline average.
 Each pink dot represents the average March temperature for a given year, and the solid blue line indicates a 5-year moving average to highlight underlying climate trends.
@@
 ```{python}
+#| label: fig-modern-temps-hist
 #| message: false
@@
 #| warning: false
-#| fig-cap: WRITE SOMETHING
+#| fig-align: center
+
+# histogram of temperatures
+plt.figure(figsize=(12, 7))
+plt.hist(modern['avg_temp_march_c'], bins=30, color='#FF9ECD', edgecolor='#E75480', alpha=0.7)
+plt.title('Distribution of Average March Temperatures (Modern)', fontsize=18, color='#E75480')
+plt.xlabel('Temperature (°C)', fontsize=14, color='#4BA3C3')
+plt.ylabel('Frequency', fontsize=14, color='#4BA3C3')
+plt.xticks(color='#4F86C6')
+plt.yticks(color='#4F86C6')
+plt.legend(framealpha=0.7)
+plt.grid(True, alpha=0.3, color='#E0E0E0')
+plt.tight_layout()
+```
+
+@fig-modern-temps-hist above presents the frequency distribution of average March temperatures in Japan from 1950 to 2020.
+The temperature values are plotted on the x-axis in degrees Celsius (°C).
+The distribution is centered around -1.5°C, with a wider spread toward colder extremes and a thinner tail extending into positive temperatures.
+While most March temperatures during this period fall below freezing, the presence of warmer years, particularly in the latter decades, indicates a shift in climate norms.
+These warming trends align with observed shifts in cherry blossom phenology, where earlier flowering is increasingly correlated with milder spring conditions.
+
+```{python}
+#| label: fig-modern-flowering-days-hist
+#| message: false
+#| echo: false
+#| warning: false
 #| fig-align: center
@@
 plt.grid(True, alpha=0.3, color='#E0E0E0')
-plt.gca().set_facecolor('#F8F8FF')  # Light background
-plt.gcf().set_facecolor('#F8F8FF')
 plt.tight_layout()
@@
 ```
-This histogram highlights the frequency distribution of sakura full flowering dates in modern Japan, measured in day-of-year (DOY), for the period between 1950 and 2020.
+
+@fig-modern-flowering-days-hist highlights the frequency distribution of sakura full flowering dates in modern Japan, measured in day-of-year (DOY), for the period between 1950 and 2020.
 Each bar represents how often flowering occurred within a specific DOY range.
-The distribution is roughly bell-shaped and unimodal, with the majority of flowering dates clustered between DOY 83 and DOY 95 (approximately March 24 to April 4), suggesting a central tendency in late March.
+The distribution is roughly normally distributed, with the majority of flowering dates clustered between DOY 83 and DOY 95 (approximately March 24 to April 4), suggesting a central tendency in late March.
 
-```{python}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: WRITE SOMETHING
-#| fig-align: center
-
-# histogram of temperatures
-plt.figure(figsize=(12, 7))
-plt.hist(modern['avg_temp_march_c'], bins=30, color='#FF9ECD', edgecolor='#E75480', alpha=0.7)
-plt.title('Distribution of Average March Temperatures (Modern)', fontsize=18, color='#E75480')
-plt.xlabel('Temperature (°C)', fontsize=14, color='#4BA3C3')
-plt.ylabel('Frequency', fontsize=14, color='#4BA3C3')
-plt.xticks(color='#4F86C6')
-plt.yticks(color='#4F86C6')
-plt.legend(framealpha=0.7)
-plt.grid(True, alpha=0.3, color='#E0E0E0')
-plt.gca().set_facecolor('#F8F8FF')  # Light background
-plt.gcf().set_facecolor('#F8F8FF')
-plt.tight_layout()
-plt.show()
-```
-The figure above presents the frequency distribution of average March temperatures in Japan from 1950 to 2020.
-The temperature values are plotted on the x-axis in degrees Celsius (°C).
-The distribution is centered around -1.5°C, with a wider spread toward colder extremes and a thinner tail extending into positive temperatures.
-While most March temperatures during this period fall below freezing, the presence of warmer years, particularly in the latter decades, indicates a shift in climate norms.
-These warming trends align with observed shifts in cherry blossom phenology, where earlier flowering is increasingly correlated with milder spring conditions.
-
-## Data Exploration
+Exploring the data collected in the last 60 years has highlighted it's similarity to the  historical dataset.
+As temperatures have gotten warmer, cherry blossoms in Kyoto have tended to bloom earlier and earlier as the years have progressed.
+Although the temperatures in the modern dataset tend to be a bit cooler on average than the ones noted in the historical dataset, the clear negative relationship between temperature and blooming day endures.
+To be able to more exactly quantify this relationship, however, we construct a model.
 
 # Modeling {#sec-model}
-
-asl;dkfj;laksdfjl;asjdf;lasjdfljasl;fj;lasjdf
 
@@
 
+## Ecological Implications
+
+## Cultural Implications
+
 ## Limitations
+
 We would like to address several methodological limiting factors.
@@
 A more comprehensive multivariate model would potentially yield enhanced predictive capacity and ecological insights.
-
-## Ecological Implications
-
-## Cultural Implications
-
-## Limitations
 
@@
 
+## Conclusion
+The Japanese tradition of *hanami*  is an important cultural tradition that honors the cycle of life and death.
+However, rising global temperatures have created an anomaly in recent years, highlighting the need to understand the historical and modern links between temperature and cherry blossom bloom timing.
+By exploring and modelling this data, we began to understand the quantitative relationships between temperature and blooming day, which allowed us to come to several conclusions.
+First, changes in temperature will have consequential effects for wildlife that function around the timing of blooming *sakura*.
+Next, the timing of celebrations such as *hanami* will be impacted by temperature changes, necessitating flexibility in honouring this tradition.
+Finally, there is a need for more a more comprehensive inquiry into the determinants of cherry blossom blooming, beyond solely the locale of Kyoto.
+This is so that more robust predictions of *sakura* blooming in Kyoto can be made.
+By gathering more data and make better predictions, the city of Kyoto will better be able to weather climate change and bridge their cherished traditions into the future.
+
 # References

### 2025-04-03 paper/paper.qmd
Update paper.qmd
committed model basic discussions and equation. also included code that uses pickled OLS models for faster visualization rendering.

@@
 from datetime import datetime
+import pickle
+import os
 
@@
 
+After exploring our data, we first considered a wide variety of models to engage with before settling on using an OLS model to best interpret the data.
+
+Initially, we considered a Cox Proportional Hazards model, which was inspired by similar applications of a Bayesian Survival model in @Allen2014.
+While this model seemed promising at first, there were a variety of challenges.
+While this model excels in handling censored data, it is less appropriate for datasets with structurally missing or entirely unrecorded entries, which characterized a large proportion of our historical data.
+Furthermore, the interpretive framework of the Cox Regression, which was centered around hazard ratios and survival probabilities, did not necessarily align with the explanatory goals we had in mind for this paper.
+
+We also evaluated the possibility of using a Poisson regression model.
+However, the dependent variable (the day of year of sakura blooming) exhibited overdispersion during modelling, and was also normally distributed.
+This violates many of the core assumptions required to produce a proper interpretation from the Poisson model, rendering it statistically unsuitable for our purposes.
+
+Ultimately, we narrowed between two candidates: the Ordinary Least Squares regression and a Regression Tree model.
+Given the structure of our dataset and the importance of interpretability, we selected the OLS model as our primary analytical tool.
+It’s coefficient-based output allowed us to quantify, interpret, and properly follow the effect of temperature and blooming times in a straightforward
+
 ## Model Specification
 
+We are interested in investigating the relationship between sakura blooming times and the average temperature in March of the given year.
+The independent variable will be that of the temperature, with the dependent/responding variable being the time of blooming.
+
+The aim of computing this OLS regression model is to estimate an equation for a line with the coefficient B1 so that the difference (error) between the predicted line and the data points are minimized.
+By doing this, we get an equation for a line that best fits the data, allowing for the estimation of blooming dates given the average March temperatures in a given year.
+With the OLS regression, we will determine and predict one single outcome variable.
+
+The equation for our linear model is written as:
+
+\begin{center}
+\text{Blooming\_DoY}_i = \beta_0 + \beta_1 \cdot \text{March\_Temp}_i + \varepsilon_i
+\end{center}
+
+The dependent variable Blooming Day (DOY) represents the blooming days of Sakuras in the given year, which is measured from 0 to 365 to denote how many days of the year has passed until blooming.
+
+The independent variable, which interacts with the slope coefficient β₁, March Temperature denotes the March Temperature of the given year, which is measured in celsius.
+
 ## Results
+
+```{python}
+with open("model/ols.pkl", "rb") as ols:
+    model_0 = pickle.load(ols)
+
+with open(model/ols_robust.pkl", "rb") as ols_robust:
+    model_1 = pickle.load(ols_robust)
+```
 

### 2025-04-03 paper/paper.qmd
refined author names in references and updated abstract for clarity and detail

--- 2025-04-03 paper/paper.qmd
@@
 abstract:
-    This is the abstract of your paper. Write a brief summary of your research,
-    including your objectives, methods, key results, and main conclusions.
-    Keep it concise but informative.
+    This paper examines the relationship between temperature and cherry blossom (sakura) flowering dates in Kyoto, Japan, analyzing a comprehensive dataset spanning nearly 1200 years from 812 CE to present day.
+While previous research has focused on shorter timeframes, this analysis integrates historical records compiled by Aono (2010) with contemporary Japanese Meteorological Agency data to better understand how climate change affects the timing of this culturally significant phenomenon.
+By modeling flowering day as a function of mean March temperatures, we identify significant trends that have implications for the traditional practice of hanami (flower-viewing) and local ecosystems.
+Our findings demonstrate the need for cultural adaptation to shifting bloom schedules and highlight the importance of extensive historical climate data in developing accurate predictive models for natural phenomena affected by global warming.
 format:
@@
 plt.ylabel('Day of Year (DOY)')
+plt.legend()
 plt.grid(True, alpha=0.3)
-plt.legend()
 plt.tight_layout()

### 2025-04-03 paper/paper.qmd

@@
 
-@fig-modern-year-temp above presents the annual full flowering dates of cherry blossoms in Japan from 1950 to 2020, measured in day-of-year (DOY), where higher values represent later blooming.
+@fig-modern-year-temp presents the annual full flowering dates of cherry blossoms in Japan from 1950 to 2020, measured in day-of-year (DOY), where higher values represent later blooming.
 Each pink dot marks the full flowering day for a given year, while the solid blue line shows a 5-year moving average, smoothing short-term fluctuations to reveal broader temporal trends.
@@
 
-@fig-modern-temps-hist above presents the frequency distribution of average March temperatures in Japan from 1950 to 2020.
+@fig-modern-temps-hist presents the frequency distribution of average March temperatures in Japan from 1950 to 2020.
 The temperature values are plotted on the x-axis in degrees Celsius (°C).
@@
 This temperature-dependent pattern aligns with the documented global warming trend observed over recent centuries, particularly the accelerated temperature increase in the modern era.
+The predictive models generated through our research demonstrate robust statistical evidence that the progressive elevation of mean March temperatures, driven by global warming, will likely continue to advance cherry blossom flowering dates in coming decades.
+Our findings corroborate the projections established by Allen et al.
+in their seminal work on phenological responses to climate change [@Allen2014]; their model, as discussed in our literature review, similarly predicted accelerated advancement of flowering dates under various warming scenarios, with comparable directional trends to those observed in our analysis.
+This convergence of independent methodological approaches strengthens the validity of both models and provides additional confirmation of the phenological impacts of climate alteration on various sakura species.
 
-The predictive models generated through our research demonstrate robust statistical evidence that the progressive elevation of mean March temperatures, driven by global warming, will likely continue to advance cherry blossom flowering dates in coming decades.
-This phenological shift carries significant ecological, cultural, and economic implications.
-The alteration of these biorhythms may disrupt established ecological synchronizations between cherry trees and their pollinators, potentially affecting ecosystem functionality and biodiversity dynamics.
+## Ecological Implications
+The observed advancement in sakura flowering dates correlates with increasing mean March temperatures in our analysis, highlighting a significant climate-driven phenological shift.
+This temperature-phenology relationship aligns with a growing body of research on plant responses to climate change.
+As @Rahimi2025 observed in their comprehensive analysis of pollinator-dependent crops across Asia, climate change is driving significant spatial and temporal shifts in biological systems, particularly affecting species interactions that depend on synchronized timing.
 
+The earlier flowering of sakura represents a classic example of phenological mismatch—where interacting species respond differently to climate cues, potentially disrupting ecological relationships.
+@Gérard2020 highlight that "many phenological shifts have already been reported at the species level for the insect emergence or blooming time of flowering plants", with temperature being the predominant driver.
+However, these shifts are rarely uniform across different ecological levels, creating potential desynchronization between plants and their pollinators or other dependent species.
+
+The observed advancement in sakura blooming could have cascading effects throughout Japanese ecosystems.
+As @Rahimi2025 identified in their spatial mismatch analysis, when flowering phenology changes, it can disrupt the "critical synchronizations" between flowers and their respective pollinators, potentially affecting ecosystem functionality and biodiversity dynamics.
+In addition, not limited to spatial mismatch, Rahimi and Jung argue that temporal mismatch is also possible in the future.
+It will occur when the "flowering periods of plants and the activity periods of their pollinators are no longer synchronized", reducing pollination efficiency and crop yields.
+This phenomenon will be a likely future if global warming continues to worsen and average March temperature steadily increases.
+
+The continued warming projected in climate models suggests that the trend toward earlier sakura flowering will likely accelerate in coming decades, potentially creating novel ecological communities as species respond differentially to changing conditions.
+
+## Cultural Implications
 From a cultural perspective, the temporal shift in cherry blossom seasons need adjustments to traditional hanami celebrations and tourism schedules, which have historically been calibrated to conventional blooming periods.
@@
 
-## Ecological Implications
-
-## Cultural Implications
-
 ## Limitations
-
 We would like to address several methodological limiting factors.

### 2025-04-03 paper/paper.qmd
fixed minor citation misspelling

@@
 The earlier flowering of sakura represents a classic example of phenological mismatch—where interacting species respond differently to climate cues, potentially disrupting ecological relationships.
-@Gérard2020 highlight that "many phenological shifts have already been reported at the species level for the insect emergence or blooming time of flowering plants", with temperature being the predominant driver.
+@Gerard2020 highlight that "many phenological shifts have already been reported at the species level for the insect emergence or blooming time of flowering plants", with temperature being the predominant driver.
 However, these shifts are rarely uniform across different ecological levels, creating potential desynchronization between plants and their pollinators or other dependent species.

### 2025-04-03 paper/paper.qmd
Update paper.qmd
added some vizes and pickles

@@
 ```{python}
-with open("model/ols.pkl", "rb") as ols:
+## Load the Pickled Models for faster visualizations
+with open("../model/ols.pkl", "rb") as ols:
     model_0 = pickle.load(ols)
 
-with open(model/ols_robust.pkl", "rb") as ols_robust:
+with open("../model/ols_robust.pkl", "rb") as ols_robust:
     model_1 = pickle.load(ols_robust)
+```
+
+```{python}
+## Print nonrobust model summary
+print(model_0.summary())
+```
+
+```{python}
+## Print robust HC1 model summary
+print(model_1.summary())
+```
+
+```{python}
+## Visualizations
+# Scatterplot with regression line
+plt.figure(figsize=(10, 6))
+sns.regplot(x='march_temp_c', y='flower_doy', data=df, ci=None, line_kws={'color': 'red'})
+plt.text(0.1, 95, f"blooming_day = {model_0.params[1]:.2f} * temp + {model_0.params[0]:.2f}", color="red", fontsize=12)
+plt.title("Sakura Blooming Day vs March Temperature")
+plt.xlabel("March Avg Temperature (°C)")
+plt.ylabel("Blooming Day of Year (DoY)")
+plt.show()
+```
+
+```{python}
+# Scatterplot with year-based color gradient
+plt.figure(figsize=(10, 6))
+scatter = plt.scatter(df['march_temp_c'], df['flower_doy'], c=df['year'], cmap='viridis', s=60, edgecolor='k')
+plt.plot(X['march_temp_c'], model_0.predict(X), color='red', label='OLS Fit')
+plt.text(-0.2, 95, f"blooming_day = {model_0.params[1]:.2f} * temp + {model_0.params[0]:.2f}", color="red", fontsize=12)
+plt.colorbar(scatter, label='Year')
+plt.title("Sakura Blooming DoY vs March Temperature with Year Gradient")
+plt.xlabel("March Avg Temperature (°C)")
+plt.ylabel("Blooming Day of Year (DoY)")
+plt.legend()
+plt.show()
 ```

### 2025-04-04 paper/paper.qmd
Remove unused model files and update file paths in Jupyter notebook

--- 2025-04-03 paper/paper.qmd
@@
 import os
+from statsmodels.iolib.summary2 import summary_col
 
@@
 
-@fig-modern-year-temp presents the annual full flowering dates of cherry blossoms in Japan from 1950 to 2020, measured in day-of-year (DOY), where higher values represent later blooming.
+@fig-modern-year-temp above presents the annual full flowering dates of cherry blossoms in Japan from 1950 to 2020, measured in day-of-year (DOY), where higher values represent later blooming.
 Each pink dot marks the full flowering day for a given year, while the solid blue line shows a 5-year moving average, smoothing short-term fluctuations to reveal broader temporal trends.
@@
 
-@fig-modern-temps-hist presents the frequency distribution of average March temperatures in Japan from 1950 to 2020.
+@fig-modern-temps-hist above presents the frequency distribution of average March temperatures in Japan from 1950 to 2020.
 The temperature values are plotted on the x-axis in degrees Celsius (°C).
@@
 
-After exploring our data, we first considered a wide variety of models to engage with before settling on using an OLS model to best interpret the data.
+After data exploration, various modeling approaches were considered before selecting OLS regression as the most appropriate method.
+We were interested in investigating the relationship between sakura blooming times and the average temperature in March of the given year.
+The independent variable will be that of the temperature, with the dependent/responding variable being the time of blooming.
 
-Initially, we considered a Cox Proportional Hazards model, which was inspired by similar applications of a Bayesian Survival model in @Allen2014.
-While this model seemed promising at first, there were a variety of challenges.
-While this model excels in handling censored data, it is less appropriate for datasets with structurally missing or entirely unrecorded entries, which characterized a large proportion of our historical data.
-Furthermore, the interpretive framework of the Cox Regression, which was centered around hazard ratios and survival probabilities, did not necessarily align with the explanatory goals we had in mind for this paper.
+A Cox Proportional Hazards model was initially evaluated, inspired by the Bayesian Survival model used in @Allen2014.
+This approach was deemed unsuitable due to its emphasis on censored data rather than structurally missing entries which characterized much of the historical dataset.
+Additionally, the interpretive framework of hazard ratios and survival probabilities was not aligned with the explanatory objectives of this analysis.
 
-We also evaluated the possibility of using a Poisson regression model.
-However, the dependent variable (the day of year of sakura blooming) exhibited overdispersion during modelling, and was also normally distributed.
-This violates many of the core assumptions required to produce a proper interpretation from the Poisson model, rendering it statistically unsuitable for our purposes.
+Poisson regression was also assessed but rejected due to overdispersion in the dependent variable (day of year of sakura blooming) and its normal distribution, which violated key assumptions required for proper model interpretation.
 
-Ultimately, we narrowed between two candidates: the Ordinary Least Squares regression and a Regression Tree model.
-Given the structure of our dataset and the importance of interpretability, we selected the OLS model as our primary analytical tool.
-It’s coefficient-based output allowed us to quantify, interpret, and properly follow the effect of temperature and blooming times in a straightforward
+The final consideration was between Ordinary Least Squares regression and Regression Tree models.
+OLS regression was ultimately selected based on the dataset structure and need for interpretability.
+Its coefficient-based output allowed for straightforward quantification and interpretation of the relationship between temperature and blooming times.
 
@@
 
-We are interested in investigating the relationship between sakura blooming times and the average temperature in March of the given year.
-The independent variable will be that of the temperature, with the dependent/responding variable being the time of blooming.
-
-The aim of computing this OLS regression model is to estimate an equation for a line with the coefficient B1 so that the difference (error) between the predicted line and the data points are minimized.
+The aim of computing this OLS regression model is to estimate an equation for a line with the coefficients β so that the differences (errors) between the predicted line and the data points are minimized.
 By doing this, we get an equation for a line that best fits the data, allowing for the estimation of blooming dates given the average March temperatures in a given year.
-With the OLS regression, we will determine and predict one single outcome variable.
+With the OLS regression, we will determine and predict one single outcome variable, blooming day.
 
@@
 
-\begin{center}
+$$
 \text{Blooming\_DoY}_i = \beta_0 + \beta_1 \cdot \text{March\_Temp}_i + \varepsilon_i
-\end{center}
+$$
 
-The dependent variable Blooming Day (DOY) represents the blooming days of Sakuras in the given year, which is measured from 0 to 365 to denote how many days of the year has passed until blooming.
+The dependent variable Blooming Day represents the blooming day of a Kyoto cherry blossoms in a given year. It is measured in days after January 1 of a given year.
 
-The independent variable, which interacts with the slope coefficient β₁, March Temperature denotes the March Temperature of the given year, which is measured in celsius.
+The independent variable March Temperature, described by the slope coefficient $\beta_1$, denotes the mean March temperature of the given year, which is measured in celsius.
+
+\newpage
 
@@
 ```{python}
-with open("model/ols.pkl", "rb") as ols:
-    model_0 = pickle.load(ols)
+# Load dataset
+df = modern_bloom
 
-with open(model/ols_robust.pkl", "rb") as ols_robust:
-    model_1 = pickle.load(ols_robust)
+# Rename columns for clarity
+df.columns = ['year', 'flower_date', 'flower_doy', 'march_temp_c']
+
+# Split into training (1953–2000) and testing (2010–2019)
+train_df = df[df['year'] <= 2000]
+test_df = df[(df['year'] >= 2009) & (df['year'] <= 2019)]
+
+# Define X and y
+X = sm.add_constant(train_df['march_temp_c'])  # Add intercept
+y = train_df['flower_doy']
+
+# Fit model 1
+model_0 = sm.OLS(y, X).fit()
+
+# Fit model 2 (with robust standard errors)
+model_1 = sm.OLS(y, X).fit(cov_type='HC1')
+
+```
+
+```{python}
+#| label: tbl-reg-results
+#| tbl-cap: "Regression Results for Cherry Blossom Flowering Day vs. March Temperature"
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Extract model results in a cleaner format
+results_table = summary_col([model_0, model_1], 
+                          model_names=["Standard OLS", "Robust OLS"],
+                          float_format='%0.3f',
+                          info_dict={'N':lambda x: "{0:d}".format(int(x.nobs)),
+                                    'R2':lambda x: "{:.3f}".format(x.rsquared)})
+# Display the table
+display(results_table)
+```
+
+The regression analysis in @tbl-reg-results reveals a significant inverse relationship between March temperature and cherry blossom blooming dates.
+For each 1°C increase in March temperature, cherry blossoms bloom approximately 2.25 days earlier (p < 0.01).
+This relationship remains consistent across both standard and robust OLS models, though the latter shows larger standard errors, suggesting heteroscedasticity in the data.
+With an R-squared of 0.32, March temperature appears to explain about one-third of the variation in blooming dates.
+When March temperature is 0°C, blossoms are predicted to emerge on the 88th day of the year (late March).
+These findings indicate that spring temperature serves as a meaningful predictor of cherry blossom phenology.
+
+```{python}
+#| label: fig-reg-plt 
+#| fig-cap: "Regression Visualization"
+
+# Simple scatterplot with regression line
+plt.figure(figsize=(8, 4))
+plt.scatter(df['march_temp_c'], df['flower_doy'], color='#FF9ECD', edgecolor='#E75480', alpha=0.7)
+plt.plot(sorted(df['march_temp_c']), model_0.predict(sm.add_constant(sorted(df['march_temp_c']))), color='red')
+plt.text(0.1, 95, f"blooming_day = {model_0.params[1]:.2f} * temp + {model_0.params[0]:.2f}", color="red", fontsize=12)
+plt.title("Sakura Blooming Day vs March Temperature")
+plt.xlabel("March Avg Temperature (°C)")
+plt.ylabel("Blooming Day of Year")
+plt.grid(True, alpha=0.3)
 ```
@@
 This temperature-dependent pattern aligns with the documented global warming trend observed over recent centuries, particularly the accelerated temperature increase in the modern era.
+
 The predictive models generated through our research demonstrate robust statistical evidence that the progressive elevation of mean March temperatures, driven by global warming, will likely continue to advance cherry blossom flowering dates in coming decades.
-Our findings corroborate the projections established by Allen et al.
-in their seminal work on phenological responses to climate change [@Allen2014]; their model, as discussed in our literature review, similarly predicted accelerated advancement of flowering dates under various warming scenarios, with comparable directional trends to those observed in our analysis.
-This convergence of independent methodological approaches strengthens the validity of both models and provides additional confirmation of the phenological impacts of climate alteration on various sakura species.
+This phenological shift carries significant ecological, cultural, and economic implications.
+The alteration of these biorhythms may disrupt established ecological synchronizations between cherry trees and their pollinators, potentially affecting ecosystem functionality and biodiversity dynamics.
 
-## Ecological Implications
-The observed advancement in sakura flowering dates correlates with increasing mean March temperatures in our analysis, highlighting a significant climate-driven phenological shift.
-This temperature-phenology relationship aligns with a growing body of research on plant responses to climate change.
-As @Rahimi2025 observed in their comprehensive analysis of pollinator-dependent crops across Asia, climate change is driving significant spatial and temporal shifts in biological systems, particularly affecting species interactions that depend on synchronized timing.
-
-The earlier flowering of sakura represents a classic example of phenological mismatch—where interacting species respond differently to climate cues, potentially disrupting ecological relationships.
-@Gerard2020 highlight that "many phenological shifts have already been reported at the species level for the insect emergence or blooming time of flowering plants", with temperature being the predominant driver.
-However, these shifts are rarely uniform across different ecological levels, creating potential desynchronization between plants and their pollinators or other dependent species.
-
-The observed advancement in sakura blooming could have cascading effects throughout Japanese ecosystems.
-As @Rahimi2025 identified in their spatial mismatch analysis, when flowering phenology changes, it can disrupt the "critical synchronizations" between flowers and their respective pollinators, potentially affecting ecosystem functionality and biodiversity dynamics.
-In addition, not limited to spatial mismatch, Rahimi and Jung argue that temporal mismatch is also possible in the future.
-It will occur when the "flowering periods of plants and the activity periods of their pollinators are no longer synchronized", reducing pollination efficiency and crop yields.
-This phenomenon will be a likely future if global warming continues to worsen and average March temperature steadily increases.
-
-The continued warming projected in climate models suggests that the trend toward earlier sakura flowering will likely accelerate in coming decades, potentially creating novel ecological communities as species respond differentially to changing conditions.
-
-## Cultural Implications
 From a cultural perspective, the temporal shift in cherry blossom seasons need adjustments to traditional hanami celebrations and tourism schedules, which have historically been calibrated to conventional blooming periods.
@@
 
+\newpage
+
 # Appendix {#sec-appendix}

### 2025-04-04 paper/paper.qmd
Add new references and enhance discussion on climate-driven phenological shifts in sakura blooming

@@
 plt.xlabel("March Avg Temperature (°C)")
-plt.ylabel("Blooming Day of Year")
-plt.grid(True, alpha=0.3)
+plt.ylabel("Blooming Day of Year (DoY)")
+plt.show()
+```
+
+```{python}
+# Scatterplot with year-based color gradient
+plt.figure(figsize=(10, 6))
+scatter = plt.scatter(df['march_temp_c'], df['flower_doy'], c=df['year'], cmap='viridis', s=60, edgecolor='k')
+plt.plot(X['march_temp_c'], model_0.predict(X), color='red', label='OLS Fit')
+plt.text(-0.2, 95, f"blooming_day = {model_0.params[1]:.2f} * temp + {model_0.params[0]:.2f}", color="red", fontsize=12)
+plt.colorbar(scatter, label='Year')
+plt.title("Sakura Blooming DoY vs March Temperature with Year Gradient")
+plt.xlabel("March Avg Temperature (°C)")
+plt.ylabel("Blooming Day of Year (DoY)")
+plt.legend()
+plt.show()
 ```
@@
 This temperature-dependent pattern aligns with the documented global warming trend observed over recent centuries, particularly the accelerated temperature increase in the modern era.
+The predictive models generated through our research demonstrate robust statistical evidence that the progressive elevation of mean March temperatures, driven by global warming, will likely continue to advance cherry blossom flowering dates in coming decades.
+Our findings corroborate the projections established by Allen et al.
+in their seminal work on phenological responses to climate change [@Allen2014]; their model, as discussed in our literature review, similarly predicted accelerated advancement of flowering dates under various warming scenarios, with comparable directional trends to those observed in our analysis.
+This convergence of independent methodological approaches strengthens the validity of both models and provides additional confirmation of the phenological impacts of climate alteration on various sakura species.
 
-The predictive models generated through our research demonstrate robust statistical evidence that the progressive elevation of mean March temperatures, driven by global warming, will likely continue to advance cherry blossom flowering dates in coming decades.
-This phenological shift carries significant ecological, cultural, and economic implications.
-The alteration of these biorhythms may disrupt established ecological synchronizations between cherry trees and their pollinators, potentially affecting ecosystem functionality and biodiversity dynamics.
+## Ecological Implications
+The observed advancement in sakura flowering dates correlates with increasing mean March temperatures in our analysis, highlighting a significant climate-driven phenological shift.
+This temperature-phenology relationship aligns with a growing body of research on plant responses to climate change.
+As @Rahimi2025 observed in their comprehensive analysis of pollinator-dependent crops across Asia, climate change is driving significant spatial and temporal shifts in biological systems, particularly affecting species interactions that depend on synchronized timing.
 
+The earlier flowering of sakura represents a classic example of phenological mismatch—where interacting species respond differently to climate cues, potentially disrupting ecological relationships.
+@Gerard2020 highlight that "many phenological shifts have already been reported at the species level for the insect emergence or blooming time of flowering plants", with temperature being the predominant driver.
+However, these shifts are rarely uniform across different ecological levels, creating potential desynchronization between plants and their pollinators or other dependent species.
+
+The observed advancement in sakura blooming could have cascading effects throughout Japanese ecosystems.
+As @Rahimi2025 identified in their spatial mismatch analysis, when flowering phenology changes, it can disrupt the "critical synchronizations" between flowers and their respective pollinators, potentially affecting ecosystem functionality and biodiversity dynamics.
+In addition, not limited to spatial mismatch, Rahimi and Jung argue that temporal mismatch is also possible in the future.
+It will occur when the "flowering periods of plants and the activity periods of their pollinators are no longer synchronized", reducing pollination efficiency and crop yields.
+This phenomenon will be a likely future if global warming continues to worsen and average March temperature steadily increases.
+
+The continued warming projected in climate models suggests that the trend toward earlier sakura flowering will likely accelerate in coming decades, potentially creating novel ecological communities as species respond differentially to changing conditions.
+
+## Cultural Implications
 From a cultural perspective, the temporal shift in cherry blossom seasons need adjustments to traditional hanami celebrations and tourism schedules, which have historically been calibrated to conventional blooming periods.

### 2025-04-04 paper/paper.qmd
Refactor code structure and improve readability in Jupyter notebooks

@@
 
-The relationship between sakura blooming date and other factors was first examined in depth.
+The Japanese tradition of Hanami has become a cultural juggernaut for the nation, even spreading globally.
+Its cultural significance cannot be understated, which underscores the importance of being able to predict when the blooming season for cherry blossoms will begin.
+The cultural significance of Hanami is not only a historic and cultural cornerstone, but has also become an economic one.
+Around the 8th century, sakura flowering had begun to assume special importance.
+Anthropologists note that in early Japanese agrarian society, cherry blossoms were closely tied to rice farming – the bloom of the cherries was believed to forecast the year’s rice harvest @CITEHanyan.
+Festivals would be conducted at shrines and the imperial court, where the emperor would offer prayers under blooming cherries for a bountiful autumn harvest @CITEHanyan.
+Sakura began to symbolize agrarian prosperity and “life power,” intimately linking Japanese livelihoods with seasonal rhythms of blooming @CITEHanyan.
+Furthermore, the blossoms started to be seen as extensions of imperial authority – one analysis notes that the emperor’s spiritual responsibility for the nation “extended ‘as far as cherries are in bloom’” @CITEHanyan, demonstrating how the flower was initially imbued with the cultural significance it holds now.
+
+Cherry blossoms continued to ingrain themselves into Japanese culture.
+While Hanami was originally an aristocratic pastime, initiatives by Japanese leaders evolved into a widespread tradition shared by all classes lup.lub.lu.se.
+Historical records indicate feudal lords and government officials would encourage their subjects to participate in planting trees as well as the viewing of their bloom lup.lub.lu.se.
+Hanami gradually became widespread, and had begun to influence Japanese art and culture substantially.
+An exhibit and the Library of Congress reveals the extent to this influence, showing through several eras and lords that the cherry blossom remained a common motif to symbolize seasonal joy and life’s fleeting pleasures for the Japanese people @CITEcongress.
+There was also religious significance to the cherry blossom, due to the influence of Buddhism infused the practice of Hanami as a reminder of life’s impermanence @CITEHanyan.
+Thus, by the early modern period, the cultural symbolism of sakura had acquired multiple layers: on one hand, it represented renewal, beauty, and communal celebration, and on the other, a poignant reminder of mortality @CITEHanyan – regardless of its artistic symbolism, the widespread appreciation of the cherry blossom bloom had cemented the season as a cultural cornerstone.
+
+This leads us to current day Japan.
+Sakura season is now a major economic and touristic phenomenon.
+The season is promoted by travel agencies, rail companies, and local tourism boards both domestically and internationally.
+Famous cherry blossom spots are must-see destinations each spring.
+As noted in @CITEjapantimes - “Hanami...
+has been a nationwide event for a long time in Japan.
+The beautiful sakura has now become a tourism asset, attracting foreign tourists”.
+Data highlights the sheer economic scale of the season: the estimated the economic effect of hanami in 2024 was ≈$7.7 billion USD in Japan, nearly double the previous year’s figure @CITEjapantimes.
+The economic impact of the cherry blossom is in no small part due to its cultural and historical significance in Japanese culture.
+This brings us to the importance of modelling this data.
+The surge of visitors to parks and landmarks during peak bloom sometimes creates logistical challenges, as vendors and tourist destinations must prepare for the influx of visitors domestic and international.
+Without adequate time to prepare for Hanami, locales risk being flooded which would cause a disruption to the revenue that Japan has come to expect @CITEmoriuchi.
+
+To summarize, the tradition of Hanami in Japan can be considered timeless.
+It’s roots in agarian traditions has also cemented a practice of recording the data around cherry blossom blooming, due to its ties to the harvest and perceived prosperity.
+It’s gradual establishment in Japanese art and culture has also made in a juggernaut in that regard, cementing its place as a staple in Japanese culture.
+Despite no longer needing to track the blossoming for the harvest, a new necessity for the data collection has arisen: economic benefit.
+With ever changing circumstances due to climate change, accurately predicting the Hanami period is more important than ever, as the sheer scale of the event for Japan culturally and economically requires substantial preparation for the celebrations to run smoothly.
+Otherwise, Japan could risk losing reliable revenue or potentially damage its reputation as a global cultural powerhouse.
+
+The relationship between sakura blooming date and other factors was next examined in depth.
 A notable amount of scholarship addresses the relationship between different variables and and the flowering dates of sakura, from humanistic cultural studies perspectives ranging to scientific plant biology, phenological modeling, and climatology studies.
@@
 
+\newpage
+
 # References
+
+::: {#refs}
+:::
 
@@
 
+\appendix
+
 # Appendix {#sec-appendix}

### 2025-04-04 paper/paper.qmd
committing prediction and diagnostic models to qmd.qmd
added more model visualizations

@@
 
+```{python}
+#| label: fig-train-plt
+#| fig-cap: "Trained Predictions Visualization"
+# Test OLS model (nonrobust)
+model_test = model_0
+
+# Prepare test data
+X_test = sm.add_constant(test_df['march_temp_c'])
+y_test = test_df['flower_doy']
+
+# Predict on test data
+y_pred = model_test.predict(X_test)
+
+# Evaluation metrics
+rmse = np.sqrt(mean_squared_error(y_test, y_pred))
+mae = mean_absolute_error(y_test, y_pred)
+r2 = r2_score(y_test, y_pred)
+
+# Plot actual vs predicted
+plt.figure(figsize=(10, 6))
+plt.plot(test_df['year'], y_test, label='Actual Blooming DoY', marker='o')
+plt.plot(test_df['year'], y_pred, label='Predicted Blooming DoY', marker='x')
+plt.xlabel("Year")
+plt.ylabel("Blooming Day of Year")
+plt.title("Actual vs Predicted Sakura Blooming (2001–2019)")
+plt.legend()
+plt.grid(True)
+plt.show()
+
+# Evaluation metrics print
+print("Evaluation on Test Set (2001–2019):")
+print(f"RMSE: {rmse:.2f}")
+print(f"MAE:  {mae:.2f}")
+print(f"R²:   {r2:.3f}")
+```
+
+```{python}
+#| label: fig-diag-results
+#| fig-cap: "Diagnostic Results for Cherry Blossom Flowering Day vs. March Temperature"
+
+# Create diagnostic plots for the OLS model
+fig = plt.figure(figsize=(20, 15))
+
+# Residual vs Fitted Plot
+ax1 = fig.add_subplot(2, 2, 1)
+residuals = model_0.resid
+fitted = model_0.fittedvalues
+sns.residplot(x=fitted, y=residuals, lowess=True, scatter_kws={'alpha': 0.5}, line_kws={'color': 'red', 'lw': 1}, ax=ax1)
+ax1.set_title('Residuals vs Fitted')
+ax1.set_xlabel('Fitted values')
+ax1.set_ylabel('Residuals')
+ax1.axhline(y=0, color='k', linestyle='-', alpha=0.3)
+
+# QQ Plot for Normality
+ax2 = fig.add_subplot(2, 2, 2)
+qqplot(model_0.resid, line='s', ax=ax2)
+ax2.set_title('Normal Q-Q')
+ax2.set_xlabel('Theoretical Quantiles')
+ax2.set_ylabel('Standardized Residuals')
+
+# Scale-Location Plot (Square root of absolute residuals vs fitted)
+ax3 = fig.add_subplot(2, 2, 3)
+sqrt_abs_resid = np.sqrt(np.abs(residuals))
+sns.regplot(x=fitted, y=sqrt_abs_resid, lowess=True, scatter_kws={'alpha': 0.5}, line_kws={'color': 'red', 'lw': 1}, ax=ax3)
+ax3.set_title('Scale-Location')
+ax3.set_xlabel('Fitted values')
+ax3.set_ylabel('$\\sqrt{|Standardized Residuals|}$')
+
+plt.tight_layout()
+
+# Leverage vs. Standardized Residuals
+ax4 = fig.add_subplot(2, 2, 4)
+infl = model_0.get_influence()
+hatvals = infl.hat_matrix_diag
+stud_resid = infl.resid_studentized_external
+
+# Determine leverage threshold
+p = len(model_0.params)  # number of parameters
+leverage_threshold = 2 * p / len(hatvals)
+
+# Plot points
+ax4.scatter(hatvals, stud_resid, alpha=0.6)
+ax4.axhline(y=0, color='k', linestyle='-', alpha=0.3)
+
+# Add reference lines
+ax4.axhline(y=2, color='g', linestyle='--', label='Std. Residuals = 2')
+ax4.axhline(y=-2, color='g', linestyle='--')
+ax4.axvline(x=leverage_threshold, color='b', linestyle='--', 
+           label=f'Leverage threshold ({leverage_threshold:.3f})')
+
+# Highlight and label influential points (leverage > threshold or std. residuals > 2)
+influential_points = []
+for i in range(len(hatvals)):
+    if (abs(stud_resid[i]) > 2) or (hatvals[i] > leverage_threshold):
+        ax4.text(hatvals[i], stud_resid[i], 
+                f" {i} ({hatvals[i]:.3f}, {stud_resid[i]:.2f})", 
+                fontsize=9, color="red")
+        influential_points.append(i)
+
+ax4.set_title('Leverage vs. Standardized Residuals')
+ax4.set_xlabel('Leverage (hat values)')
+ax4.set_ylabel('Studentized Residuals')
+ax4.legend(loc='upper right', fontsize=8)
+
+plt.tight_layout()
+plt.show()
+```
+
 # Discussion {#sec-discussion}

### 2025-04-04 paper/paper.qmd
Refine conclusion section for clarity and conciseness in discussion of cherry blossom phenology

--- 2025-04-04 paper/paper.qmd
@@
 However, rising global temperatures have created an anomaly in recent years, highlighting the need to understand the historical and modern links between temperature and cherry blossom bloom timing.
-By exploring and modelling this data, we began to understand the quantitative relationships between temperature and blooming day, which allowed us to come to several conclusions.
-First, changes in temperature will have consequential effects for wildlife that function around the timing of blooming *sakura*.
+By exploring and modelling this data, we began to understand the quantitative relationships between temperature and blooming day.
+Changes in temperature will have consequential effects for wildlife that function around the timing of blooming *sakura*.
 Next, the timing of celebrations such as *hanami* will be impacted by temperature changes, necessitating flexibility in honouring this tradition.
@@
 This is so that more robust predictions of *sakura* blooming in Kyoto can be made.
-By gathering more data and make better predictions, the city of Kyoto will better be able to weather climate change and bridge their cherished traditions into the future.
+By gathering more data to make better predictions, the city of Kyoto will be better able to weather climate change and bridge their cherished traditions into the future.
 

### 2025-04-04 paper/paper.qmd
Revise README to enhance project overview and clarify data structure for sakura flowering analysis

--- 2025-04-04 paper/paper.qmd
@@
 
+\newpage
+
 The dependent variable Blooming Day represents the blooming day of a Kyoto cherry blossoms in a given year. It is measured in days after January 1 of a given year.
@@
 The independent variable March Temperature, described by the slope coefficient $\beta_1$, denotes the mean March temperature of the given year, which is measured in celsius.
-
-\newpage
 
@@
 These findings indicate that spring temperature serves as a meaningful predictor of cherry blossom phenology.
+Diagnostic plots are in @sec-appendix-1
 
@@
 #| fig-cap: "Regression Visualization"
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
 
@@
 ```{python}
-# Scatterplot with year-based color gradient
+#| label: fig-train-plt
+#| fig-cap: "Trained Predictions Visualization"
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Test OLS model (nonrobust)
+
+model_test = model_0
+
+# Prepare test data
+X_test = sm.add_constant(test_df['march_temp_c'])
+y_test = test_df['flower_doy']
+
+# Predict on test data
+y_pred = model_test.predict(X_test)
+
+# Plot actual vs predicted
 plt.figure(figsize=(10, 6))
-scatter = plt.scatter(df['march_temp_c'], df['flower_doy'], c=df['year'], cmap='viridis', s=60, edgecolor='k')
-plt.plot(X['march_temp_c'], model_0.predict(X), color='red', label='OLS Fit')
-plt.text(-0.2, 95, f"blooming_day = {model_0.params[1]:.2f} * temp + {model_0.params[0]:.2f}", color="red", fontsize=12)
-plt.colorbar(scatter, label='Year')
-plt.title("Sakura Blooming DoY vs March Temperature with Year Gradient")
-plt.xlabel("March Avg Temperature (°C)")
-plt.ylabel("Blooming Day of Year (DoY)")
+plt.plot(test_df['year'], y_test, label='Actual Blooming DoY', marker='o')
+plt.plot(test_df['year'], y_pred, label='Predicted Blooming DoY', marker='x')
+plt.xlabel("Year")
+plt.ylabel("Blooming Day of Year")
+plt.title("Actual vs Predicted Sakura Blooming (2001–2019)")
 plt.legend()
-plt.show()
+plt.grid(True)
 ```
+
+\newpage
 
@@
 ## Cultural Implications
-From a cultural perspective, the temporal shift in cherry blossom seasons need adjustments to traditional hanami celebrations and tourism schedules, which have historically been calibrated to conventional blooming periods.
-In addition, these modifications could impact cultural practices that have remained relatively consistent for centuries and negatively impact tourism economy that is built around it.
+As established in @sec-lit-review, that tradition of Hanami is a staple in Japanese culture.
+Moreover, it has become a significant source of revenue for both the nation itself and the local communities which host associated festivals.
+Our findings reveal a significant relationship between pre-flowering temperatures and blooming dates, aligning closely with observed patterns with climate-change related temperature increases.
+As the Hanami/cherry blossom festivals draw significant international and domestic attention, the scheduling of these events must be exhaustive in order to facilitate them at the scale they reach (and continue to grow).
+While we don’t anticipate a significant disruption to the bloom itself, understanding when the bloom will start is critical to maintaining the economic momentum these festivals have demonstrated.
+Furthermore, cherry blossoms and the Hanami practice are significant beyond their economic benefits.
+Communities and citizens celebrate and find great meaning and importance in these practices.
+Understanding that the bloom may come earlier, and being able to plan for travel, accommodations, and individual scheduling is important to many people in Japan.
+We expect that, if temperatures continue to raise year over year, it may disrupt the lives/scheduling of communities in Japan which find great importance in observing cherry blossom blooming every year.
 
 ## Limitations
-We would like to address several methodological limiting factors.
-Firstly, our analysis relies predominantly on data from Kyoto, representing a geographically limited subset that may not comprehensively reflect sakura phenological patterns across Japan's diverse climate regions.
-Secondly, although we used a merged dataset spanning from 800 CE to 2019, there are significant data gaps/missing values where we have to drop during the cleaning & merging of the two dataset.
+There are several limitations to this analysis.
+First, our analysis relies predominantly on data from Kyoto, representing a geographically limited subset that may not comprehensively reflect sakura phenological patterns across Japan's diverse climate regions.
+Second, although we used a merged dataset spanning from 800 CE to 2019, there are significant data gaps/missing values where we have to drop during the cleaning & merging of the two dataset.
 In addition, a notable limitation is the historical dataset's reliability, as large portions comprise of reconstructed values derived from local linear regression models or originated from primary sources that lack rigorous academic verification protocols.
@@
 Variables such as precipitation patterns, photoperiod fluctuations, soil conditions, and winter chilling requirements may also contribute significantly to the observed advancement in flowering dates.
-A more comprehensive multivariate model would potentially yield enhanced predictive capacity and ecological insights.
+A more comprehensive multivariate model, comprising more reliable data would potentially yield enhanced predictive capacity and ecological insights.
 
@@
 # Appendix {#sec-appendix}
+
+## Diagnostic Plots for OLS Models {#sec-appendix-1}
+
+```{python}
+#| fig-cap: "Diagnostic Results for Cherry Blossom Flowering Day vs. March Temperature"
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Add necessary libraries for diagnostic plots
+import statsmodels.api as sm
+from statsmodels.graphics.gofplots import qqplot
+import matplotlib.pyplot as plt
+import seaborn as sns
+import numpy as np
+
+# Create diagnostic plots for the OLS model
+fig = plt.figure(figsize=(20, 15))
+
+# Residual vs Fitted Plot
+ax = fig.add_subplot(1, 1, 1)
+residuals = model_0.resid
+fitted = model_0.fittedvalues
+sns.residplot(x=fitted, y=residuals, lowess=True, scatter_kws={'alpha': 0.5}, line_kws={'color': 'red', 'lw': 1}, ax=ax)
+ax.set_title('Residuals vs Fitted')
+ax.set_xlabel('Fitted values')
+ax.set_ylabel('Residuals')
+ax.axhline(y=0, color='k', linestyle='-', alpha=0.3)
+```
+
+```{python}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+import statsmodels.api as sm
+from statsmodels.graphics.gofplots import qqplot
+import matplotlib.pyplot as plt
+import seaborn as sns
+import numpy as np
+
+# Create diagnostic plots for the OLS model
+fig = plt.figure(figsize=(10, 8))
+ax = plt.subplot(111)
+
+# QQ Plot for Normality
+qqplot(model_0.resid, line='s', ax=ax)
+ax.set_title('Normal Q-Q')
+ax.set_xlabel('Theoretical Quantiles')
+ax.set_ylabel('Standardized Residuals')
+```
+
+```{python}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Add necessary libraries for diagnostic plots
+import statsmodels.api as sm
+from statsmodels.graphics.gofplots import qqplot
+import matplotlib.pyplot as plt
+import seaborn as sns
+import numpy as np
+
+# Create diagnostic plots for the OLS model
+fig = plt.figure(figsize=(10, 8))
+
+# Scale-Location Plot (Square root of absolute residuals vs fitted)
+ax = fig.add_subplot(2, 2, 3)
+sqrt_abs_resid = np.sqrt(np.abs(residuals))
+sns.regplot(x=fitted, y=sqrt_abs_resid, lowess=True, scatter_kws={'alpha': 0.5}, line_kws={'color': 'red', 'lw': 1}, ax=ax)
+ax.set_title('Scale-Location')
+ax.set_xlabel('Fitted values')
+ax.set_ylabel('$\\sqrt{|Standardized Residuals|}$')
+
+```
+
+```{python}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Add necessary libraries for diagnostic plots
+import statsmodels.api as sm
+from statsmodels.graphics.gofplots import qqplot
+import matplotlib.pyplot as plt
+import seaborn as sns
+import numpy as np
+
+# Leverage vs. Standardized Residuals
+fig = plt.figure(figsize=(10, 8))
+ax = fig.add_subplot(1, 1, 1)
+infl = model_0.get_influence()
+hatvals = infl.hat_matrix_diag
+stud_resid = infl.resid_studentized_external
+
+# Determine leverage threshold
+p = len(model_0.params)  # number of parameters
+leverage_threshold = 2 * p / len(hatvals)
+
+# Plot points
+ax.scatter(hatvals, stud_resid, alpha=0.6)
+ax.axhline(y=0, color='k', linestyle='-', alpha=0.3)
+
+# Add reference lines
+ax.axhline(y=2, color='g', linestyle='--', label='Std. Residuals = 2')
+ax.axhline(y=-2, color='g', linestyle='--')
+ax.axvline(x=leverage_threshold, color='b', linestyle='--', 
+           label=f'Leverage threshold ({leverage_threshold:.3f})')
+
+# Highlight and label influential points (leverage > threshold or std. residuals > 2)
+influential_points = []
+for i in range(len(hatvals)):
+    if (abs(stud_resid[i]) > 2) or (hatvals[i] > leverage_threshold):
+        ax.text(hatvals[i], stud_resid[i], 
+                f" {i} ({hatvals[i]:.3f}, {stud_resid[i]:.2f})", 
+                fontsize=9, color="red")
+        influential_points.append(i)
+
+ax.set_title('Leverage vs. Standardized Residuals')
+ax.set_xlabel('Leverage (hat values)')
+ax.set_ylabel('Studentized Residuals')
+ax.legend(loc='upper right', fontsize=8)
+plt.tight_layout()
+```

### 2025-04-04 paper/paper.qmd
finalest final commit

@@
 
-The regression analysis in @tbl-reg-results reveals a significant inverse relationship between March temperature and cherry blossom blooming dates.
+The regression analysis in @tbl-reg-results, visualized in @fig-reg-plt, reveals a significant inverse relationship between March temperature and cherry blossom blooming dates.
 For each 1°C increase in March temperature, cherry blossoms bloom approximately 2.25 days earlier (p < 0.01).
@@
 plt.scatter(df['march_temp_c'], df['flower_doy'], color='#FF9ECD', edgecolor='#E75480', alpha=0.7)
+
 plt.plot(sorted(df['march_temp_c']), model_0.predict(sm.add_constant(sorted(df['march_temp_c']))), color='red')
+
 plt.text(0.1, 95, f"blooming_day = {model_0.params[1]:.2f} * temp + {model_0.params[0]:.2f}", color="red", fontsize=12)
+
 plt.title("Sakura Blooming Day vs March Temperature")
@@
 plt.show()
-```
-
-```{python}
-#| label: fig-train-plt
-#| fig-cap: "Trained Predictions Visualization"
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-align: center
-
-# Test OLS model (nonrobust)
-
-model_test = model_0
-
-# Prepare test data
-X_test = sm.add_constant(test_df['march_temp_c'])
-y_test = test_df['flower_doy']
-
-# Predict on test data
-y_pred = model_test.predict(X_test)
-
-# Plot actual vs predicted
-plt.figure(figsize=(10, 6))
-plt.plot(test_df['year'], y_test, label='Actual Blooming DoY', marker='o')
-plt.plot(test_df['year'], y_pred, label='Predicted Blooming DoY', marker='x')
-plt.xlabel("Year")
-plt.ylabel("Blooming Day of Year")
-plt.title("Actual vs Predicted Sakura Blooming (2001–2019)")
-plt.legend()
-plt.grid(True)
 ```
@@
 # Appendix {#sec-appendix}
+
+## Code for Plots
+
+The code used to render the plots and tables in this paper can be found at <https://github.com/lcarnegie/sakura/tree/main/scripts>
 

### 2025-04-26 paper/paper.qmd
re-rendered paper to make nicer

@@
 Our findings demonstrate the need for cultural adaptation to shifting bloom schedules and highlight the importance of extensive historical climate data in developing accurate predictive models for natural phenomena affected by global warming.
-format:
-    pdf:
-        documentclass: article
-        papersize: letter
-        toc: true
-        number-sections: true
-        colorlinks: true
-        fig-cap-location: bottom
-        tbl-cap-location: top
-        fontsize: "12"
-        geometry: "margin=1.2in" # print margins (change if wanting to)
+format: pdf
 bibliography: "references.bib"
